### 2015-06-02 Appendices/AppendixCPETRawData.tex
Latex files added for first time

% Appendix CPET Raw Data

\chapter{Breath-by-breath CPET sample data} % Main appendix title

\label{AppendixCPETRawData} % For referencing this appendix elsewhere, use \ref{AppendixA}

\lhead{Appendix C. \emph{Raw CPET Data}} % This is for the header on each page - perhaps a shortened title
\begin{figure}
	\centering
	\caption{Breath-by-breath sample data with values averaged every 10 seconds - Part 1.}
	\includegraphics[width=\textwidth,page=1]{Figures/cpet_9p_rawdata.pdf}
\end{figure}
\begin{figure}
	\centering
	\caption{Breath-by-breath sample data with values averaged every 10 seconds - Part 2.}
	\includegraphics[width=\textwidth,page=2]{Figures/cpet_9p_rawdata.pdf}
\end{figure}