### 2015-06-02 Tables/c1_clavien_dindo.tex

\begin{sidewaystable}[htbp]
\caption{The Clavien-Dindo Classification of Surgical Complications}
\label{table:clavien-dindo}
\centering
\begin{tabular}{|l m{15cm}|}
 \hline
 Grade  &   Description  \\ \hline

 Grade I     &   Any deviation from the normal postoperative course without the need for pharmacological treatment or surgical, endoscopic and radiological interventions. \\

Grade II      &   Requiring pharmacological treatment with drugs other than such allowed for grade I complications. Blood transfusions and total parenteral nutrition are also included.\\ 
Grade III     &   Requiring surgical, endoscopic or radiological intervention\\
                &  Grade III-a: - intervention not under general anesthesia\\
                &  Grade III-b: - intervention under general anesthesia\\ 
Grade IV       &  Life-threatening complication (including CNS complications)‡ requiring IC/ICU-management\\
                &   Grade IV-a: - single organ dysfunction (including dialysis)\\
                &   Grade IV-b: - multi organ dysfunction\\ 
Grade V        &   Death of a patient\\ \hline
Suffix 'd':     &   If the patients suffers from a complication at the time of discharge,  the suffix  “d”  (for ‘disability’) is added to the respective grade of complication. This label indicates the need for a follow-up to fully evaluate the complication.\\ \hline

\end{tabular}

\end{sidewaystable}