### 2015-06-02 Tables/c1_cpet_parameters.tex

\begin{table}[p]
\centering
\caption{Parameters measured at spirometry.}
\label{table:spirometry}
\begin{tabular}{| m{2cm}  m{2cm}  m{7cm} |}
	\hline
	Parameter & Units  & Description                          \\ \hline
	FVC       & litres & Forced Vital Capacity                \\
	FEV1      & litres & Forced Expiratory Volume in 1 second \\
	FEV1/FVC  & \%     & Tiffeneau-Pinelli[1] index           \\ \hline
\end{tabular}

\vspace{2cm}

\caption{Common parameters measured at cardiopulmonary exercise testing.}
\label{table:cpet_parameters}
\begin{tabular}{| m{2cm}  m{2cm}  m{7cm} |}
	\hline
	Parameter & Units       & Description                                  \\ \hline
	\%peakVO2 & \%          & VO2 as a \% of predicted VO2Peak             \\
	Load      & Watts       & Exercise Workload                            \\
	VE        & litres/min  & Ventilatory Equivalent                       \\
	Vt        & litres      & Tidal volume                                 \\
	VO2       & litres/min  & Absolute Oxygen uptake/consumption           \\
	VO2/kg    & ml/(kg*min) & Corrected Oxygen uptake/consumption          \\
	VE/VO2    &             & Ventilatory Equivalent for O$_2$             \\
	VCO2      & litres/min  & Carbon-dioxide output                        \\
	VE/VCO2   &             & Ventilatory Equivalent for CO$_2$            \\
	RER       &             & Respiratory Exchange Ratio                   \\
	PETO2     & mmHg        & End Tidal O2                                 \\
	PETCO2    & mmHg        & End Tidal CO2                                \\
	O2Pulse   & ml/beat     & Oxygen pulse                                 \\
	HR        & beats/min   & Heart Rate                                   \\
	Bf        & /min        & Breathing Frequency                          \\
	P(A-a)O2  & mmHg        & Alveolar-arterial PO2 difference             \\
	Vd/Vt     &             & Physiologic dead space-to-tidal volume ratio \\
	SBP       & mmHg        & Systolic blood pressure                      \\
	DBP       & mmHg        & Diastolic blood pressure                     \\
	O2sat     & \%          & Oxygen saturation                            \\ \hline
\end{tabular}

\end{table}