### 2015-06-02 main.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Masters/Doctoral Thesis 
% LaTeX Template
% Version 1.43 (17/5/14)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original authors:
% Steven Gunn 
% http://users.ecs.soton.ac.uk/srg/softwaretools/document/templates/
% and
% Sunil Patel
% http://www.sunilpatel.co.uk/thesis-template/
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
% Note:
% Make sure to edit document variables in the Thesis.cls file
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[12pt, oneside]{Thesis} % The default font size, sides to print and document class

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%--------------------PACKAGES------------------------------------------------------------

\usepackage[backend=biber,
			style=authoryear,
			url=false]{biblatex}
%For Square Brackets around citations
%http://tex.stackexchange.com/questions/16765/biblatex-author-year-square-brackets
\makeatletter

\newrobustcmd*{\parentexttrack}[1]{%
	\begingroup
	\blx@blxinit
	\blx@setsfcodes
	\blx@bibopenparen#1\blx@bibcloseparen
	\endgroup}

\AtEveryCite{%
	\let\parentext=\parentexttrack%
	\let\bibopenparen=\bibopenbracket%
	\let\bibcloseparen=\bibclosebracket}

\makeatother
%------------------------------------------------------

\usepackage{graphicx}
\usepackage{caption} 
\usepackage{subcaption} % For Subfigures - not sure if I have used this anywhere
\usepackage{parskip} %Adds extra space between paragraphs
\usepackage{listings} %This is for the VBA code in the appendices
\usepackage{amsmath} %For aligning all equations on the equal to sign
\usepackage{pdfpages} %for including pdf files or individual pages of pdf files
\usepackage[pagewise]{lineno} %To help Donny review. Remove for final version
\linenumbers

%--------------Table customisations------------------------------------------------------
% The following allows creation of columns of defined width within tables
%http://tex.stackexchange.com/questions/12703/how-to-create-fixed-width-table-columns-with-text-raggedright-centered-raggedlef
\usepackage{array}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

% The next few lines are for customising tables
\usepackage[justification=justified,singlelinecheck=false]{caption}
\usepackage{rotating} % for landscape tables use begin{sidewaystable}
%----------------Figures and hyperlinks----------------------------------------------------
%\graphicspath{Figures/} % Specifies the directory where pictures are stored
%\hypersetup{urlcolor=black, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%Bibligraphy file from Zotero
\addbibresource{zot.bib}
%----------------------------------------------------------------------------------------

\title{\ttitle} % Defines the thesis title - don't touch this

%----------------------------------------------------------------------------------------
% BEGIN DOCUMENT -- BEGIN DOCUMENT -- BEGIN DOCUMENT -- BEGIN DOCUMENT -- BEGIN DOCUMENT
%----------------------------------------------------------------------------------------
\begin{document}

\frontmatter % Use roman page numbering style (i, ii, iii, iv...) for the pre-content pages

\setstretch{2.0} % Line spacing of 2.0 although I have set \doublespace set somewhere else

% Define the page headers using the FancyHdr package and set up for one-sided printing
\fancyhead{} % Clears all page headers and footers
\rhead{\thepage} % Sets the right side header to show the page number
\lhead{} % Clears the left side page header

\pagestyle{fancy} % Finally, use the "fancy" page style to implement the FancyHdr headers

\newcommand{\HRule}{\rule{\linewidth}{0.5mm}} % New command to make the lines in the title page

% PDF meta-data
\hypersetup{pdftitle={\ttitle}}
\hypersetup{pdfsubject=\subjectname}
\hypersetup{pdfauthor=\authornames}
\hypersetup{pdfkeywords=\keywordnames}

%----------------------------------------------------------------------------------------
%	TITLE PAGE - Reformatted to Univ of Glasgow guidelines
%----------------------------------------------------------------------------------------

\begin{titlepage}
	\begin{center}

	\HRule \\[0.4cm] % Horizontal line
	{\huge \bfseries \ttitle}\\[0.4cm] % Thesis title
	\HRule \\[1.5cm] % Horizontal line

	\begin{minipage}{1\textwidth}
		\begin{center} \large
			\setstretch{1.3}
			\textbf{by} \\
			\vspace{1.1cm}
			\textbf{\authornames}\\
			\textbf{M.B.B.S., M.R.C.S., Dip.N.B.}\\
			\vspace{1.1cm}
			\normalsize
			\textbf{SUBMITTED IN FULFILMENT OF THE REQUIREMENTS}\\
			\textbf{FOR THE DEGREE OF DOCTOR OF MEDICINE}\\
			\vspace{1.1cm}
			\textbf{to}\\
			\vspace{1.1cm}
			\textbf{THE UNIVERSITY OF GLASGOW}\\
			\vspace{1.1cm}
			\textbf{BASED ON RESEARCH CONDUCTED IN THE} \\
			\textbf{UNIVERSITY DEPARTMENT OF SURGERY,}\\
			\textbf{GLASGOW ROYAL INFIRMARY}\\
			\textbf{AUGUST 2015}\\
			\vspace{1.1cm}
			\textcopyright \textbf{Vishnu V Chandrabalan 2015}
		\end{center}

	\end{minipage}
	\vfill
	\end{center}

\end{titlepage}

%----------------------------------------------------------------------------------------
%	QUOTATION PAGE
%----------------------------------------------------------------------------------------

\pagestyle{empty} % No headers or footers for the following pages

\dedicatory{Pebbles on the beach.}%Never again.
\clearpage % Start a new page

%----------------------------------------------------------------------------------------
%	DEDICATION
%----------------------------------------------------------------------------------------

\pagestyle{empty} % Page style needs to be empty for this page

\dedicatory{Dedicated to A, I, A} % Dedication text

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics

%----------------------------------------------------------------------------------------
%	ABSTRACT PAGE
%----------------------------------------------------------------------------------------

\addtotoc{Abstract} % Add the "Abstract" page entry to the Contents

\abstract{\addtocontents{toc}{\vspace{1em}} % Add a gap in the Contents, for aesthetics

To be finalised...
}
\clearpage

%----------------------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES PAGES
%----------------------------------------------------------------------------------------

\pagestyle{fancy} % The page style headers have been "empty" all this time, now use the "fancy" headers as defined before to bring them back

\lhead{\emph{Contents}} % Set the left side page header to "Contents"
\tableofcontents % Write out the Table of Contents

\lhead{\emph{List of Figures}} % Set the left side page header to "List of Figures"
\listoffigures % Write out the List of Figures

\lhead{\emph{List of Tables}} % Set the left side page header to "List of Tables"
\listoftables % Write out the List of Tables

\clearpage

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

\acknowledgements{\addtocontents{toc}{\vspace{1em}} % Add a gap in the Contents, for aesthetics

I would like to thank the following people, for their help, advice and encouragement:

Professor Donald C McMillan,\\
University Department of Surgery, Glasgow Royal Infirmary

Professor Paul Horgan,\\
University Department of Surgery, Glasgow Royal Infirmary
}

\clearpage % Start a new page

%----------------------------------------------------------------------------------------
%	DECLARATION PAGE
%	Your institution may give you a different text to place here
%----------------------------------------------------------------------------------------

\Declaration{

	\addtocontents{toc}{\vspace{1em}} % Add a gap in the Contents, for aesthetics

	I declare that the work presented in this thesis was carried out solely by me, as a clinical research fellow in the University Dept of Surgery, Royal Infirmary, Glasgow, except where indicated below:

	Measurement of biochemical and haematological data was performed by the hospital laboratory service.

	Statistical analysis was performed with the assistance of Prof Donald C McMillan, University Dept of Surgery, Royal Infirmary, Glasgow.

	In addition, no work referred to in this thesis has been submitted in support of an application for another degree or qualification in this or any other university.
}

%----------------------------------------------------------------------------------------
%	ABBREVIATIONS
%----------------------------------------------------------------------------------------

\input{Chapters/Abbreviations}

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - CHAPTERS
%----------------------------------------------------------------------------------------

\mainmatter % Begin numeric (1,2,3...) page numbering

\pagestyle{fancy} % Return the page headers back to the "fancy" style

% Include the chapters of the thesis as separate files from the Chapters folder
\input{Chapters/Introduction}
\input{Chapters/CPET_Outcomes} 
\input{Chapters/CPET_Jaundice}
\input{Chapters/BodyComp}
\input{Chapters/CPET_Survival}
\input{Chapters/Discussion}

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - APPENDICES
%----------------------------------------------------------------------------------------

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics

\appendix % Cue to tell LaTeX that the following 'chapters' are Appendices

%\input{Appendices/AppendixExcelMacros}
\input{Appendices/AppendixCPETRawData}
%\input{Appendices/AppendixC}

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics

\backmatter

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

\label{Bibliography}

\lhead{\emph{Bibliography}} % Change the page header to say "Bibliography"

%\bibliographystyle{unsrtnat} % Use the "unsrtnat" BibTeX style for formatting the Bibliography
\printbibliography % The references (bibliography) information are stored in the file named "Bibliography.bib"

\end{document}