### 2015-07-10 Tables/crp_comp_ROC_infections_noPOPF.tex
CRP_Complications - Minor Changes
- Multiple minor cosmetic changes

%07/07/15 - Started this table
\begin{table}[h]
	\centering
	\caption{Receiver operating characteristics curve analysis of C-reactive protein as a marker of postoperative infective complications in patients who did not develop a POPF}
	\label{table:crp_comp_ROC_infections_noPOPF}
	\renewcommand{\arraystretch}{1.2} %Increases space between rows
	%\setlength{\tabcolsep}{9pt} %sets the space between columns
	\begin{tabular}{| c | c c c | c c c c c |}
		\hline
		Day & AUC  & p        & 95\% CI     & CRP & Spec. & Sens. & NPV  & PPV               \\ \hline
		3   & 0.64 & 0.011    & 0.54 - 0.74 & 178 & 0.66  & 0.64  & 0.79 & 0.48              \\
		4   & 0.71 & $<$0.001 & 0.61 - 0.80 & 125 & 0.64  & 0.74  & 0.83 & 0.50              \\
		5   & 0.70 & $<$0.001 & 0.61 - 0.80 & 107 & 0.64  & 0.63  & 0.78 & 0.45              \\
		6   & 0.73 & $<$0.001 & 0.65 - 0.82 & 92  & 0.68  & 0.73  & 0.84 & 0.53              \\
		7   & 0.80 & $<$0.001 & 0.72 - 0.88 & 87  & 0.65  & 0.81  & 0.87 & 0.52              \\ \hline
		\multicolumn{9}{l}{AUC - Area under curve, Spec. - Specificity, Sens. - Sensitivity} \\
		\multicolumn{9}{l}{NPV - Negative Predictive Value, PPV - Positive Predictive Value}
	\end{tabular}
\end{table}