### 2015-07-10 Tables/crp_comp_vs_POPF_ISGPS_p_values_only.tex

%06/07/15 - Started this table
\begin{table}[h]
	\centering
	\caption{The relationship between CRP during the first postoperative week and post-operative pancreatic fistula (POPF) in patients undergoing pancreaticoduodenectomy.}
	\label{table:crp_comp_vs_POPF_ISGPS_p_values_only}
	\renewcommand{\arraystretch}{1.2} %Increases space between rows
	%\setlength{\tabcolsep}{9pt} %sets the space between columns
	\begin{tabular}{| c | c c c c |}
		\hline
		           &       \multicolumn{4}{c}{POPF ISGPS Grades - \textit{p}}        \\
		Postop Day & No vs. Any Grade & No vs. A & A vs. B & B vs. C                 \\ \hline
		0          & 0.722            & 0.582    & 0.849   & 0.730                   \\
		1          & 0.242            & 0.331    & 0.740   & 0.295                   \\
		2          & 0.002            & 0.161    & 0.222   & 0.563                   \\
		3          & $<$0.001         & 0.022    & 0.602   & 0.530                   \\
		4          & $<$0.001         & 0.007    & 0.906   & 0.834                   \\
		5          & $<$0.001         & 0.035    & 0.611   & 0.959                   \\
		6          & $<$0.001         & 0.012    & 0.638   & 0.781                   \\
		7          & $<$0.001         & 0.006    & 0.235   & 0.356                   \\ \hline
		\multicolumn{5}{l}{\textit{p} - Mann-Whitney U test}                         \\
		\multicolumn{5}{l}{ISGPS - International Study Group for Pancreatic Surgery}	\\
				\multicolumn{5}{l}{POPF - Postoperative Pancreatic Fistula}
	\end{tabular}

\end{table}