### 2015-08-09 main.tex
Major changes as below
- Thesis summary completed
- Many changes to thesis.cls and main.tex to format abstract page and university url, etc
- Update Intro-body comp - measuring body comp
- Zot.bin backed up and updated
- Conclusion chapter title renamed to Discussion
- It is 01:40 am - Have listened to Dr. Dre, Snoop Dogg and Eminem all day
- 2 shots of Jura and I am knackered yet sleepless after 2 strong coffees

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Masters/Doctoral Thesis 
% LaTeX Template
% Version 1.43 (17/5/14)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original authors:
% Steven Gunn 
% http://users.ecs.soton.ac.uk/srg/softwaretools/document/templates/
% and
% Sunil Patel
% http://www.sunilpatel.co.uk/thesis-template/
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
% Note:
% Make sure to edit document variables in the Thesis.cls file
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[12pt, oneside]{Thesis} % The default font size, sides to print and document class

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%--------------------PACKAGES------------------------------------------------------------

\usepackage[backend=biber,
			citestyle=authoryear,
			bibstyle=authoryear,
			uniquelist=false,
			uniquename=false,
			maxcitenames=1,
			mincitenames=1,
			url=false]{biblatex}
%For Square Brackets around citations
%http://tex.stackexchange.com/questions/16765/biblatex-author-year-square-brackets
%http://ctan.mirrorcatalogs.com/macros/latex/contrib/biblatex/doc/biblatex.pdf - for info on all the above options. File is also stored in the Glasgow Research folder
\makeatletter

\newrobustcmd*{\parentexttrack}[1]{%
	\begingroup
	\blx@blxinit
	\blx@setsfcodes
	\blx@bibopenparen#1\blx@bibcloseparen
	\endgroup}

\AtEveryCite{%
	\let\parentext=\parentexttrack%
	\let\bibopenparen=\bibopenbracket%
	\let\bibcloseparen=\bibclosebracket}

\makeatother
%------------------------------------------------------

\usepackage{graphicx}
\usepackage{caption} 
\usepackage{subcaption} % For Subfigures - not sure if I have used this anywhere
\usepackage{parskip} %Adds extra space between paragraphs
\usepackage{listings} %This is for the VBA code in the appendices
\usepackage{amsmath} %For aligning all equations on the equal to sign
\usepackage{pdfpages} %for including pdf files or individual pages of pdf files
\usepackage{easy-todo} %Adds todos to the document. Use [disable] as package option in final verion. Also see \listoftodos below.
\usepackage[pagewise]{lineno} %To help Donny review. Remove for final version
\linenumbers

%--------------Table customisations------------------------------------------------------
% The following allows creation of columns of defined width within tables
%http://tex.stackexchange.com/questions/12703/how-to-create-fixed-width-table-columns-with-text-raggedright-centered-raggedlef
\usepackage{array}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

% The next few lines are for customising tables
\usepackage[justification=justified,singlelinecheck=false]{caption}
\usepackage{rotating} % for landscape tables use begin{sidewaystable}
%----------------Figures and hyperlinks----------------------------------------------------
%\graphicspath{Figures/} % Specifies the directory where pictures are stored
%\hypersetup{urlcolor=black, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%Bibligraphy file from Zotero
\addbibresource{zot.bib}
%----------------------------------------------------------------------------------------

\title{\ttitle} % Defines the thesis title - don't touch this

%----------------------------------------------------------------------------------------
% BEGIN DOCUMENT -- BEGIN DOCUMENT -- BEGIN DOCUMENT -- BEGIN DOCUMENT -- BEGIN DOCUMENT
%----------------------------------------------------------------------------------------
\begin{document}

\frontmatter % Use roman page numbering style (i, ii, iii, iv...) for the pre-content pages

\setstretch{1.5} % Line spacing of 1.5 although I have set \doublespace set somewhere else - not sure where

% Define the page headers using the FancyHdr package and set up for one-sided printing
\fancyhead{} % Clears all page headers and footers
\rhead{\thepage} % Sets the right side header to show the page number
\lhead{} % Clears the left side page header

\pagestyle{fancy} % Finally, use the "fancy" page style to implement the FancyHdr headers

\newcommand{\HRule}{\rule{\linewidth}{0.5mm}} % New command to make the lines in the title page

% PDF meta-data
\hypersetup{pdftitle={\ttitle}}
\hypersetup{pdfsubject=\subjectname}
\hypersetup{pdfauthor=\authornames}
\hypersetup{pdfkeywords=\keywordnames}

%----------------------------------------------------------------------------------------
%	TITLE PAGE - Reformatted to Univ of Glasgow guidelines
%----------------------------------------------------------------------------------------

\begin{titlepage}
	\begin{center}

	\HRule \\[0.4cm] % Horizontal line
	{\huge \bfseries \ttitle}\\[0.4cm] % Thesis title
	\HRule \\[1.5cm] % Horizontal line

	\begin{minipage}{1\textwidth}
		\begin{center} \large
			\setstretch{1.3}
			\textbf{by} \\
			\vspace{1.1cm}
			\textbf{\authornames}\\
			\textbf{M.B.B.S., M.R.C.S., Dip.N.B.}\\
			\vspace{1.1cm}
			\normalsize
			\textbf{SUBMITTED IN FULFILMENT OF THE REQUIREMENTS}\\
			\textbf{FOR THE DEGREE OF DOCTOR OF MEDICINE}\\
			\vspace{1.1cm}
			\textbf{to}\\
			\vspace{1.1cm}
			\textbf{THE UNIVERSITY OF GLASGOW}\\
			\vspace{1.1cm}
			\textbf{BASED ON RESEARCH CONDUCTED IN THE} \\
			\textbf{UNIVERSITY DEPARTMENT OF SURGERY,}\\
			\textbf{GLASGOW ROYAL INFIRMARY}\\
			\textbf{AUGUST 2015}\\
			\vspace{1.1cm}
			\textcopyright \textbf{Vishnu V Chandrabalan 2015}
		\end{center}

	\end{minipage}
	\vfill
	\end{center}

\end{titlepage}

%----------------------------------------------------------------------------------------
%	QUOTATION PAGE
%----------------------------------------------------------------------------------------

\pagestyle{empty} % No headers or footers for the following pages

\dedicatory{Pebbles on the beach.}%Never again.
\clearpage % Start a new page

%----------------------------------------------------------------------------------------
%	DEDICATION
%----------------------------------------------------------------------------------------

\pagestyle{empty} % Page style needs to be empty for this page

\dedicatory{Dedicated to A, I, A} % Dedication text

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics

%----------------------------------------------------------------------------------------
%	ABSTRACT PAGE
%----------------------------------------------------------------------------------------

\addtotoc{Abstract} % Add the "Abstract" page entry to the Contents

\abstract{\addtocontents{toc}{\vspace{1em}} % Add a gap in the Contents, for aesthetics
Pancreaticoduodenectomy with or without adjuvant chemotherapy remains the only modality of possible cure in patients with cancer involving the head of the pancreas and the periampullary region. While mortality rates after pancreaticoduodenectomy have improved considerably over the course of the last century, morbidity remains high. Patient selection is of paramount importance in ensuring that major surgery is offered to individuals who will most benefit from a pancreaticoduodenectomy. Moreover, identifying preoperative risk factors provides potential targets for prehabilitation and optimisation of the patient's physiology before undertaking surgery. In addition to this, early identification of patients who are likely to develop postoperative complications allows for better allocation of critical care resources and more targeted treatment of the high risk patients.

Cardiopulmonary exercise testing is becoming an increasingly popular tool in the preoperative risk assessment of the surgical patient. However, very little work has been done to investigate the role of cardiopulmonary exercise test in patients scheduled to undergo pancreaticoduodenectomy. The impact of jaundice, sytemic inflammation and other preoperative clinicopathological characteristics of the patient on their cardiopulmonary exercise physiology has not been studied in detail before in this cohort of patients. Moreover, the impact of these factors on long-term survival is also unknown.

The overall aim of the thesis was to evaluate the relationship between cardiopulmonary exercise testing, perioperative clinico-pathological characteristics and postoperative outcomes as well as long-term survival in patients undergoing pancreaticoduodenectomy.

\textbf{Chapter \ref{ch_intro}} reviews the existing literature on preoperative cardiopulmonary exercise testing, the impact of obstructive jaundice, perioperative systemic inflammation and the importance of body composition in determining outcomes in patients undergoing major surgery with particular reference to pancreatic surgery.

\textbf{Chapter \ref{ch_cpet_outcomes}} reports on the role of cardiopulmonary exercise testing in predicting postoperative complications. The results demonstrate that patients with anaerobic threshold less than 10 mls/kg/min were more likely to develop a postoperative pancreatic fistula, stay longer in hospital and less likely to receive adjuvant therapy after pancreaticoduodenectomy. These results emphasise the importance of excellent aerobic fitness to improve postoperative outcomes and provide a potential target for future prehabilitation programmes.

\textbf{Chapter \ref{ch_cpet_jaundice}} evaluates the relationship between cardiopulmonary exercise physiology and other clinicopathological characteristics of the patient. In a detailed analysis of cardiopulmonary exercise test parameters in jaundiced versus non-jaundiced patients, the results in this chapter demonstrate that obstructive jaundice does not impair cardiopulmonary exercise physiology. This further supports emerging evidence in contemporary literature that jaundiced patients can proceed directly to surgery without underoing preoperative biliary drainage. The results of this study also show an interesting inverse relationship between body mass index and anaerobic threshold which is analysed in more detail in Chapter 3.

\textbf{Chapter \ref{ch_bodycomp}} examines the relationship between preoperative cardiopulmonary exercise physiology and body composition in depth. All parameters measured at cardiopulmonary exercise test are compared against body composition and body mass index. The results of this chapter report that the current method of reporting $\dot{V}_{O_2}$, both at peak exercise and anaerobic threshold, is biased against obese subjects and advises caution in the interpretation of cardiopulmonary exercise test results in patients with a high BMI. This is particularly important as current evidence in literature suggests that postoperative outcomes in obese subjects are comparable to non-obese subjects while cardiopulmonary exercise test results are also abnormally low in this very same cohort of patients.

\textbf{Chapter \ref{ch_survival}} analyses the relationship between preoperative comorbidity including cardiopulmonary exercise test and long-term survival in patients with pancreatic ductal adenocarcinoma undergoing pancreaticoduodenectomy. The results of this chapter suggest that cardiopulmonary exercise testing parameters are not useful in predicting long-term survival. However, a composite score of preoperative physiology as measured by the POSSUM Physiology score was associated with survival.

\textbf{Chapter \ref{ch_crp_comp}} evaluates the role of the early postoperative systemic inflammatory response in predicting complications after pancreaticoduodenectomy and aims to establish clinically relevant thresholds for C-Reactive Protein for the prediction of complications. The results of this chapter demonstrate that CRP levels as early as the second postoperative day are associated with complications. While post-operative CRP was useful in the prediction of infectious complications, this was only the case in patients who did not develop a post-operative pancreatic fistula. The predictive ability of inflammatory markers for infectious complications was blunted in patients with a pancreatic fistula. In these patients, outcome was largely determined by the severity of the pancreatic fistula.

\textbf{Chapter \ref{ch_discussion}} summarises the findings of this thesis, their place in current literature and future directions. The contents of this thesis aim to add to the current body of knowledge regarding the complex pathophysiological alterations that occur in patients with lesions of the head of the pancreas with the associated obstructive jaundice, systemic inflammation and their impact on cardiopulmonary exercise physiology, postoperative complications and long-term outcomes. The results establish the value of cardiopulmonary exercise test as an objective measure of aerobic fitness while refuting a commonly held belief that obstructive jaundice significantly affected cardiovascular function. The current work paves the way for further study into evaluating the role of cardiopulmonary exercise test in prehabilitation regimens as well as the incorporation of preoperative $\dot{V}_{O_2}$AT and postoperative CRP into existing models for predicting complications after pancreaticoduodenectomy. These results will help improve patient selection and preparation for pancreaticoduodenectomy  and ultimately the short and long-term outcomes after this morbid yet necessary operation.

}
\clearpage

%----------------------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES PAGES
%----------------------------------------------------------------------------------------

\pagestyle{fancy} % The page style headers have been "empty" all this time, now use the "fancy" headers as defined before to bring them back

\lhead{\emph{Contents}} % Set the left side page header to "Contents"
\tableofcontents % Write out the Table of Contents

\clearpage
\lhead{\emph{List of Figures}} % Set the left side page header to "List of Figures"
\listoffigures % Write out the List of Figures

\clearpage
\lhead{\emph{List of Tables}} % Set the left side page header to "List of Tables"
\listoftables % Write out the List of Tables

\clearpage

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

\acknowledgements{\addtocontents{toc}{\vspace{1em}} % Add a gap in the Contents, for aesthetics

I would like to thank the following people, for their help, advice and encouragement:

Professor Donald C McMillan,\\
University Department of Surgery, Glasgow Royal Infirmary

Professor Paul Horgan,\\
University Department of Surgery, Glasgow Royal Infirmary
}

\clearpage % Start a new page

%----------------------------------------------------------------------------------------
%	DECLARATION PAGE
%	Your institution may give you a different text to place here
%----------------------------------------------------------------------------------------

\Declaration{

	\addtocontents{toc}{\vspace{1em}} % Add a gap in the Contents, for aesthetics

	I declare that the work presented in this thesis was carried out solely by me, as a clinical research fellow in the University Dept of Surgery, Royal Infirmary, Glasgow, except where indicated below:

	Measurement of biochemical and haematological data was performed by the hospital laboratory service.

	Statistical analysis was performed with the assistance of Prof Donald C McMillan, University Dept of Surgery, Royal Infirmary, Glasgow.

	In addition, no work referred to in this thesis has been submitted in support of an application for another degree or qualification in this or any other university.
}

%----------------------------------------------------------------------------------------
%	ABBREVIATIONS
%----------------------------------------------------------------------------------------

\input{Chapters/Abbreviations}

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - CHAPTERS
%----------------------------------------------------------------------------------------

\mainmatter % Begin numeric (1,2,3...) page numbering

\pagestyle{fancy} % Return the page headers back to the "fancy" style

% Include the chapters of the thesis as separate files from the Chapters folder
\input{Chapters/Introduction} 		%1
%\input{Chapters/CPET_Outcomes} 	%2
%\input{Chapters/CPET_Jaundice} 	%3
%\input{Chapters/BodyComp}			%4
%\input{Chapters/CPET_Survival}		%5
%\input{Chapters/CRP_Complications}	%6
%\input{Chapters/Discussion}		%7

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - APPENDICES
%----------------------------------------------------------------------------------------

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics

\appendix % Cue to tell LaTeX that the following 'chapters' are Appendices

%\input{Appendices/AppendixExcelMacros}
%\input{Appendices/AppendixCPETRawData}
%\input{Appendices/AppendixC}

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics

\backmatter

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

\label{Bibliography}

\lhead{\emph{Bibliography}} % Change the page header to say "Bibliography"

%\bibliographystyle{unsrtnat} % Use the "unsrtnat" BibTeX style for formatting the Bibliography
\printbibliography

\listoftodos

\end{document}