### 2015-08-18 Tables/bc_at_new_indices.tex
Body Composition - Major Changes
- Tables changed to include SAT/VAT instead of CSA
- Text changed in most places accordingly
- New tables and figures to reflect new vo2 scalings
- Next commit should finalise this chapter with some luck

\begin{sidewaystable}[p]
	\caption{The relationship between body composition, body habitus and $\dot{V}_{O_2}$AT scaled using different factors in patients undergoing major pancreatic surgery.}
	\label{table:bc_at_new_indices}
	\footnotesize
	\centering
	\renewcommand{\arraystretch}{1.5} %Increases space between rows
	%\setlength{\tabcolsep}{9pt} %sets the space between columns
	%\hline\noalign{\smallskip} %increases space after the line
	\begin{tabular}{|l|cc|cc|cc|cc|cc|cc|}
		\hline
		& \multicolumn{2}{c|}{Absolute $\dot{V}_{O_2}$} & \multicolumn{2}{c|}{$\frac{\dot{V}_{O_2}}{Weight}$} & \multicolumn{2}{c|}{$\frac{\dot{V}_{O_2}}{Height^2}$} & \multicolumn{2}{c|}{$\frac{\dot{V}_{O_2}}{BMI}$} & \multicolumn{2}{c|}{$\frac{\dot{V}_{O_2}}{Skeletal\ Muscle}$}  & \multicolumn{2}{c|}{$\frac{\dot{V}_{O_2}}{Lean\ Body\ Mass}$}\\

		Variable              & $\rho$ &    p     & $\rho$ &    p     & $\rho$ &    p     & $\rho$ &    p     & $\rho$ &    p    & $\rho$ &    p \\ \hline
		Subcut. fat (cm2)     & 0.191  &  0.089   & -0.326 &  0.003   & 0.169  &  0.132   & -0.217 &  0.051   & 0.018  &  0.872  & -0.041 & 0.716 \\
		Visceral fat (cm2)    & 0.184  &  0.103   & -0.365 &  0.001   & 0.154  &  0.168   & -0.279 &  0.012   & -0.065 &  0.563   & -0.108 & 0.339\\
		Total fat (cm2)       & 0.231  &  0.038   & -0.400 & $<$0.001 & 0.191  &  0.088   & -0.286 &  0.010   & -0.021 &  0.851   & -0.082 & 0.466\\
		Skeletal Muscle (cm2) & 0.463  & $<$0.001 & -0.078 &  0.487   & 0.329  &  0.003   & 0.105  &  0.353   & -0.429 & $<$0.001 & 0.133 & 0.237 \\
		Height                & 0.446  & $<$0.001 & 0.019  &  0.870   & 0.001  &  0.993   & 0.475  & $<$0.001 & 0.105  &  0.350   & 0.019 & 0.865\\
		Weight                & 0.545  & $<$0.001 & -0.302 &  0.006   & 0.358  &  0.001   & -0.043 &  0.703   & 0.001  &  0.995   & 0.033 & 0.770\\
		Body Mass Index       & 0.357  &  0.001   & -0.363 &  0.001   & 0.425  & $<$0.001 & -0.345 &  0.002   & -0.059 &  0.603   & 0.035 & 0.756\\ \hline
		\multicolumn{11}{l}{$\rho$ - Pearson's r adjusted for sex.}
	\end{tabular}
\end{sidewaystable}