### 2015-08-20 Appendices/AppendixAccessDatabase.tex
Minor changes - Minor edits to Bodycomp, intro, main and appendix - New appendix file for Access database

\chapter{Visual Basic for Applications For Data Collection} % Main appendix title
\label{AppendixAccessDatabase} % For referencing this appendix elsewhere, use \ref{AppendixA}
\lhead{Appendix \ref{AppendixAccessDatabase}. \emph{MS Access VBA For Data Collection}} % This is for the header on each page - perhaps a shortened title

\setstretch{1.0}
%------------------------------------------