### 2015-08-22 Chapters/CPET_Jaundice.tex
Review 1 - CPET_Jaundice
- Major rewrite of introduction and discussion
- Reads a lot better now - lets see what Donny says
- Time to sleep ?!

%CPET and Jaundice

\chapter{An investigation into the relationship between cardiopulmonary exercise testing, obstructive jaundice and preoperative pathophysiology in patients undergoing major pancreatic surgery.}

\label{ch_cpet_jaundice}
\lhead{Chapter \ref{ch_cpet_jaundice}. \emph{CPET, Jaundice and Preoperative Pathophysiology}}
\clearpage

%----------------------------------------------------------------------------------------

\section{Introduction}
Patients with tumours involving the pancreatic head or the periampullary region often present with inoperable disease. 
In the minority of patients with operable disease, resectional surgery in the form of a pancreaticoduodenectomy remains the main modality of treatment and only chance of a potential cure. 
However, major pancreatic surgery is associated with significant morbidity and mortality and is only undertaken in specialist centres. 
Patient selection, preoperative optimisation, good surgical technique and improvements in postoperative care have all contributed to a reduction in mortality \parencite{winter_1423_2006} but morbidity remains high \parencite{mann_review_2010}.
While several technical strategies have been described in recent years to minimise morbidity, these strategies are not necessarily based on a better understanding of the physiological basis of postoperative complications in these patients.

The close anatomical relationship between the distal bile duct, distal pancreatic duct, head of the pancreas and the duodenum is responsible for obstructive jaundice being the most common presenting symptom in patients with tumours affecting this region. 
Distal bile duct strictures also occur in a proportion of patients with severe chronic pancreatitis involving the pancreatic head.
In such patients, it may be impossible to distinguish inflammation from malignant disease on preoperative imaging and a proportion of these patients will require a pancreaticoduodenectomy \parencite{abraham_pancreaticoduodenectomy_2003}.
The perioperative management of the patient with obstructive jaundice is complex and management algorithms are still evolving \parencite{wang_preoperative_2008}.

Obstructive jaundice has been shown to be associated with abnormal cardiovascular physiology in several animal and human studies. 
Surgery in the jaundiced patient has been reported to be associated with adverse postoperative haemodynamic events and renal dysfunction \parencite{pain_perioperative_1985,green_systemic_1995}. 
The association between jaundice and cardiovascular physiology was reported over a hundred years ago by King and co-workers who found that injection of porcine bile pigment into dogs resulted in bradycardia, hypotension and eventually death \parencite{king_effect_1909}. 
Green and co-workers (1986) described the effects of `cholemia' in dogs that were subjected to choledochocaval anastomosis. 
The resultant myocardial depression was described by them as the `jaundiced heart' \parencite{green_jaundiced_1986} and was associated with poor myocardial response to inotropic stimulation in dogs \parencite{binah_obstructive_1985, bomzon_systemic_1986} as well as humans \parencite{lumlertgul_jaundiced_1991}.

As a result, preoperative biliary drainage (PBD) used to be advocated routinely before subjecting a patient to pancreaticoduodenectomy with the intention of reducing postoperative morbidity. 
A recent multi-center, randomised trial found that PBD was associated with increased incidence of complications after pancreaticoduodenectomy \parencite{van_der_gaag_preoperative_2010}.
However, this trial excluded patients with a bilirubin levels greater than 250 $\mu$mol/l from the study. 
A Cochrane meta-analysis of 6 studies including 520 patients also concluded that pre-operative biliary drainage was associated with increased incidence of complications and should not be used routinely in patients scheduled to undergo pancreaticoduodenectomy \parencite{fang_pre-operative_2012}.

\subsection{Aim}
The results presented in Chapter \ref{ch_cpet_outcomes} demonstrated that poor performance at cardiopulmonary exercise testing (CPET) was associated with adverse outcomes after pancreaticoduodenectomy resulting in an increased incidence of POPF and prolonged hospital stay. 
Obstructive jaundice was present in a large proportion of these patients.
The aim of the present study was to evaluate the relationship between cardiopulmonary exercise testing, obstructive jaundice (especially severe obstructive jaundice with serum bilirubin $>$ 250 $\mu$mol/l) and preoperative pathophysiology in patients undergoing major pancreatic surgery.

\clearpage

\section{Patients and Methods}

\subsection{Patients}
Patients who underwent classical or pylorus-preserving pancreaticoduodenectomy for periampullary lesions (both benign and malignant) between August 2008 and April 2013 and had undergone cardiopulmonary exercise testing as part of their preoperative work-up at the West of Scotland Pancreatic Unit, Glasgow Royal Infirmary, Glasgow were included in the study. 
Established criteria for resectability in patients with malignant disease were used as outlined in previous published work. 
Segmental or wedge resection of the portal vein or superior mesenteric vein was carried out if the lesion was otherwise resectable.

\subsection{Preoperative data}
Patient demographics, preoperative clinico-pathological characteristics including cardiorespiratory comorbidity, results of preoperative blood tests, chest x-ray, ECG and cardiopulmonary exercise tests were collected from prospectively held databases. 
The POSSUM Physiology Score was calculated based on 11 physiological parameters (cardiac disease, respiratory disease, ECG changes, pulse rate, blood pressure, haemoglobin, white cell count, serum sodium, serum potassium, serum urea and Glasgow Coma Scale) and was used as an objective score of comorbidity. 
Cardiovascular comorbidity was defined as a score of 2 or more for either the cardiac disease or ECG component of the POSSUM score. 
Respiratory comorbidity was defined as a score or 2 or more for the respiratory disease component of the POSSUM score.

\subsection{Obstructive jaundice}
Serum bilirubin levels and liver function tests were measured in all patients on the day before surgery. 
Obstructive jaundice was defined as serum bilirubin $>$ 35 $\mu$mol/litre and severe obstructive jaundice was defined as serum bilirubin $>$ 250 $\mu$mol/litre. 
These thresholds were chosen as the recent randomised controlled trial that demonstrated increased complications in patients who underwent PBD excluded patients with serum bilirubin $>$ 250 $\mu$mol/litre from the study \parencite{van_der_gaag_preoperative_2010}.
The present study aimed to evaluate preoperative pathophysiology in this particular group of patients with `severe obstructive jaundice'.
Due to referral practices at the time of this study, a proportion of patients underwent preoperative biliary interventions and drainage before referral to the West of Scotland Pancreatic Unit. 
Data regarding biliary intervention and stenting was collected where feasible.

\subsection{Cardiopulmonary exercise test}
Cardiopulmonary exercise tests were performed in the Department of Respiratory Medicine at the Glasgow Royal Infirmary using the ZAN-600 CPET suite (nSpire Health, Longmont, CO 80501, USA). 
All patients underwent standard pulmonary function tests and spirometry prior to cardiopulmonary exercise testing. 
A cycle ergometer was used to perform a symptom-limited, incremental work-load test preceded by a 3-minute rest period. 
The test was stopped when patients achieved their maximum exercise tolerance, when significant ischaemic changes occurred on ECG or for other safety reasons. 
Peak oxygen consumption achieved at this stage was defined as $\dot{V}_{O_2}$Peak. 
The $\dot{V}_{O_2}$AT was calculated using the V-slope \parencite{beaver_new_1986,sue_metabolic_1988} and ventilatory equivalents \parencite{society_ats/accp_2003} methods.
$\dot{V}_{O_2}$AT less than 10 ml/kg/min was considered to be low based on previous work by us (Chapter \ref{ch_cpet_outcomes}) as well as Ausania and co-workers \parencite{ausania_effects_2012} which has shown increased incidence of complications in patients with $\dot{V}_{O_2}$AT below this threshol. 
Oxygen consumption at peak exercise ($\dot{V}_{O_2}$Peak) was dichotomised using a cut-off of 16 ml/kg/min. 
Cardiopulmonary exercise testing methodology and calculation of $\dot{V}_{O_2}$AT is described in detail in Sections \ref{sec:cpx_method} and \ref{sec:cpx_at_method}.
A mored detailed description of cardiopulmonary exercise testing parameters described in this study are published elsewhere \parencite{balady_clinicians_2010}.

\subsection{Statistics}
Grouping of the variables was carried out using standard or previously published thresholds. 
In the absence of such thresholds, the variables were treated as continuous variables. 
Non-parametric tests were used to analyse the association between categorical and continuous variables while Chi-square tests were used to analyse the association between categorical variables. 
Univariate and multivariate binary logistic regression analysis was used to study the relationship between preoperative patient characteristics and $\dot{V}_{O_2}$AT / $\dot{V}_{O_2}$Peak. 
SPSS software (Version 17.0; SPSS Inc., Chicago, IL, USA) was used to perform statistical analysis.

\clearpage

\section{Results}

\subsection{Clinico-pathological characteristics}
One-hundred and thirty eight patients underwent pancreaticoduodenectomy with preoperative cardiopulmonary exercise testing during the study period. 
Over half of the patients were male (n=93, 67\%). 
Approximately half the number of patients were over the age of 65 (n=68, 49\%) and overweight or obese (n=69, 50\%). 
Cardiovascular comorbidity was present in 58 patients (42\%) and respiratory comorbidity was present in 12 patients (9\%). 
Fifty patients (36\%) had a history of cigarette smoking. 
The POSSUM Physiology Score was greater than 14 in 61 patients (44\%). 
Obstructive jaundice (serum bilirubin 35-250 $\mu$mol/l) was present in 32 (23\%) patients while severe obstructive jaundice (serum bilirubin $>$ 250 $\mu$mol/l) was present in 19 (14\%) patients.

The baseline demographic and clinical characteristics of non-jaundiced and jaundiced patients are shown in Table \ref{table:cpet_oj_patient}. 
A greater proportion of jaundiced patients were females compared to the non-jaundiced cohort (p=0.028) and smokers (p=0.038). 
Elevated POSSUM Physiology Score (p=0.004) and malignancy (p$<$0.001) were significantly associated with the presence of jaundice. 
However, there was no statistically significant difference in age, body mass index, cardiovascular comorbidity, respiratory comorbidity or PBD between the non-jaundiced and jaundiced patients.

\input{Tables/cpet_oj_patient}

\subsection{Obstructive jaundice vs. preoperative blood tests}
The relationship between obstructive jaundice and preoperative blood tests is shown in Table \ref{table:cpet_oj_bloods}. 
Obstructive jaundice was associated with the presence of a systemic inflammatory response as characterised by raised serum C-reactive protein levels and decrease in serum albumin levels (p$<$0.001). 
The degree of systemic inflammation was proportionate to the severity of jaundice. 
Non-jaundiced patients had a median CRP of 3.6 mg/dl (inter-quartile range 0.3 - 89) and serum albumin of 37 g/dl (IQR 18 - 46) while patients with severe obstructive jaundice had a median CRP of 13 (IQR 1.7 - 51) and a median serum albumin of 25 g/dl (IQR 18 - 33). 
Obstructive jaundice was also associated with electrolyte abnormalities including serum sodium, serum potassium and serum chloride levels. 
Jaundiced patients were more likely to be anaemic (p$<$0.001) with lower haematocrit (p$<$0.001) and lower mean corpuscular volume (p=0.001). 
However, there was no significant difference in the preoperative renal function, prothrombin time and white cell count between jaundiced and non-jaundiced patients.

\input{Tables/cpet_oj_bloods}

\subsection{Obstructive jaundice vs. pulmonary function tests}
Pulmonary function tests including forced vital capacity (FVC), forced expiratory volume in one second (FEV1) and their derived parameters (predicted FEV1, FEV1/FVC, predicted FEV1/FVC) were compared between jaundiced and non-jaundiced patients. 
There was a trend towards lower forced vital capacity with increasing severity of jaundice but this did not reach statistical significance (p=0.092). 
There was no association between the other pulmonary function tests and jaundice. (Table \ref{table:cpet_oj_pft}).

\input{Tables/cpet_oj_pft}

\subsection{Obstructive jaundice vs. CPET}
Cardiopulmonary exercise test parameters measured at the anaerobic threshold and at peak exercise were compared between jaundiced and non-jaundiced patients using the Kruskal-Wallis test. 
Obstructive jaundice was associated with lower tidal volume (p=0.017), lower absolute $\dot{V}_{O_2}$ (p=0.003), lower corrected $\dot{V}_{O_2}$ (p=0.029), lower oxygen pulse (p=0.037) and lower respiratory rate (p=0.022). 
At peak exercise, jaundice was associated with lower tidal volume (p=0.028), lower absolute $\dot{V}_{O_2}$Peak (p=0.007), lower absolute $\dot{V}_{CO_2}$ (p=0.016) and lower end-tidal $O_2$ or $PET_{O_2}$ (p=0.026). 
However, as shown in Tables \ref{table:cpet_oj_anaerobic} and \ref{table:cpet_oj_peak}, several of these associations were not linear with a trend towards poorer results between non-jaundiced and mildly jaundiced patients and better values in the severely jaundiced cohort.

\input{Tables/cpet_oj_anaerobic}
\input{Tables/cpet_oj_peak}

Scatter-plot analysis comparing serum bilirubin versus $\dot{V}_{O_2}$AT and $\dot{V}_{O_2}$Peak as continuous variables is depicted in Figure \ref{fig:cpet_oj_scatter}. 
This shows that the relationship between serum bilirubin and $\dot{V}_{O_2}$AT is weak with a $\rho^2$ value of 0.035 Pearson's $\rho$ = - 0.187, p = 0.028. 
There was no correlation between serum bilirubin and $\dot{V}_{O_2}$Peak (Pearson's $\rho$ = - 0.132, $\rho^2$ = 0.017, p = 0.123)

\begin{figure}[htbp]
	\centering
	\begin{subfigure}{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Figures/cpet_oj_scatter_at_bil}
		\caption{$\dot{V}_{O_2}$AT versus serum bilirubin}
		\label{fig:cpet_oj_scatter_at_bil}
	\end{subfigure}
	\begin{subfigure}{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Figures/cpet_oj_scatter_peak_bil}
		\caption{$\dot{V}_{O_2}$Peak versus serum bilirubin}
		\label{fig:cpet_oj_scatter_peak_bil}
	\end{subfigure}

	\caption{Pearson's correlation and scatter-plot analysis comparing serum bilirubin versus $\dot{V}_{O_2}$AT and $\dot{V}_{O_2}$Peak in patients undergoing pancreaticoduodenectomy.}
	\label{fig:cpet_oj_scatter}

\end{figure}

\subsection{Preoperative factors related to CPET}
Binary logistic regression analysis was undertaken to assess preoperative clinico-pathological patient factors associated with a low $\dot{V}_{O_2}$AT and low $\dot{V}_{O_2}$Peak. 
On univariate analysis, female sex (HR 2.74, 95\% CI 1.30-5.74, p=0.008), body mass index $>$ 25 (HR 3.09, 95\% CI 1.51-6.32, p=0.002), presence of malignancy (HR 3.59, 95\% CI 1.36-9.43, p=0.010), POSSUM Physiology Score $>$ 14 (HR 2.06, 95\% CI 1.02-4.17, p= 0.044), serum bilirubin $>$ 250 $\mu$mol/l (HR 5.66, 95\% CI 1.87-17.16, p=0.002), haemoglobin $<$ 12 g/dl (HR 2.74, 95\% CI 1.30-5.74, p=0.008) and serum C-reactive protein $>$ 10 mg/dl (HR 2.18, 95\% CI 1.06-4.51, p=0.035) were associated with $\dot{V}_{O_2}$AT $<$ 10 ml/kg/min. 
Age, cardiovascular and respiratory comorbidity, PBD, mild jaundice and serum albumin were not significantly related to a low $\dot{V}_{O_2}$AT.

The statistically significant independent variables with p$<$0.05 were entered into a backward step-wise regression model with $\dot{V}_{O_2}$AT $<$ 10 ml/kg/min as a categorical, binomial, dependent variable. 
Female sex (HR 3.75, 95\% CI 1.57-8.95 p$<$0.005), body mass index $>$ 25 (HR 3.65, 95\% CI 1.61-8.26, p$<$0.005), malignancy (HR 4.02, 95\% CI 1.33-12.16, p$<$0.05) and C-reactive protein $>$ 10 mg/dl (HR 2.98, 95\% CI 1.29-6.86, p$<$0.05) were independently related to low $\dot{V}_{O_2}$AT. 
Obstructive jaundice was not related to low $\dot{V}_{O_2}$AT on multivariate analysis. (Table \ref{table:cpet_oj_at_regression})

Binary logistic regression analysis of the relationship between patient factors and low $\dot{V}_{O_2}$Peak ($<$ 16 ml/kg/min) is shown in Table \ref{table:cpet_oj_peak_regression}. 
On both univariate and multivariate analysis, female sex, body mass index $>$ 25 and haemoglobin $<$ 12 g/dl were significantly associated with a low $\dot{V}_{O_2}$Peak.

\input{Tables/cpet_oj_at_regression}
\input{Tables/cpet_oj_peak_regression}

\clearpage

\section{Discussion}

The results of the present study demonstrate that obstructive jaundice, even when severe, is not an independent risk factor for poor preoperative cardiopulmonary exercise physiology in patients undergoing pancreaticoduodenectomy. 
The results also demonstrate that obstructive jaundice is associated with elevated preoperative systemic inflammation with the severity of systemic inflammation related to the severity of the jaundice.

%History of CPET
The use of CPET in preoperative risk prediction was first made popular over two decades ago by Older and co-workers \parencite{older_preoperative_1993}. 
Since then cardiopulmonary exercise testing has been reported to be useful in identifying high risk patients prior to major general \parencite{snowden_submaximal_2010}, pancreatic \parencite{ausania_effects_2012},[Chapter \ref{ch_cpet_outcomes}], oesophagogastric \parencite{nagamatsu_preoperative_2001} as well as vascular \parencite{carlisle_mid-term_2007} surgery. 
Cardiopulmonary exercise testing is superior to conventional measures of fitness due to the dynamic nature of the test that evaluates the adequacy of oxygen delivery to tissues under physiological stress. 
However, the factors responsible for poor aerobic capacity in surgical patients have not been adequately studied. 
The present study aimed to address this in patients undergoing major pancreatic surgery.

%History of jaundice - reason for PBD
Experimental and animal studies have shown that obstructive jaundice was associated with myocardial depression \parencite{green_jaundiced_1986}, poor myocardial response to inotropic stimulation \parencite{lumlertgul_jaundiced_1991}, impaired sympathetic baroreflex sensitivity \parencite{song_baroreflex_2009} and deranged atrial natriuretic peptide levels \parencite{pereira_increased_1994,gallardo_increased_1998}.

Historically, obstructive jaundice has also been reported to be associated with adverse haemodynamic events in patients undergoing major surgery. 
Intra-operative blood loss, postoperative hypotension, increased susceptibility to shock and renal dysfunction were all more common in patients with obstructive jaundice \parencite{dixon_factors_1983, pain_perioperative_1985, green_systemic_1995}.
A recent review noted that bile acids have a complex range of receptor mediated effects on the cardiovascular system \parencite{khurana_bile_2011}. 
Moreover, some of these effects appear to be partly reversible by biliary drainage as demonstrated by Padillo and co-workers \parencite{padillo_improved_2001}.
This increased incidence of complications resulted in routine pre-operative biliary drainage being recommended in jaundiced patients before undertaking major surgery. 
Pancreaticoduodenectomy was described by Whipple initially as a two-stage operation, with the first stage involving a cholecysto-gastrostomy aimed at relieving biliary obstruction before undertaking the resection at a later second operation \parencite{whipple_treatment_1935}.

%Reasons for not doing PBD
However, several recent studies have reported that PBD is associated with increased incidence of complications and that surgery in the jaundiced patients is safe.
Pitt and coworkers in a prospective randomised trial comparing outcomes in jaundiced patients undergoing surgery with or without PBD reported that PBD was associated with increased cost without any decrease in postoperative complications \parencite{pitt_does_1985}. 
A recent meta-analysis \parencite{sewnath_meta-analysis_2002} pooled data from 5 randomised controlled trials comparing surgery with PBD versus surgery without drainage.
The authors concluded that PBD not only did not improve postoperative complication rates or mortality but resulted in a higher overall complication rate due to the morbidity associated with the procedure itself. 
These findings have been reproduced in a more recent multi-center, randomised trial \parencite{van_der_gaag_preoperative_2010}.
A recent Cochrane Collaboration review of six trials including 520 patients concluded that PBD may be associated with serious adverse events and must not be performed routinely outwith trial settings \parencite{wang_preoperative_2008}.

%CPET vs Jaundice
In the present study, obstructive jaundice was not independently related to $\dot{V}_{O_2}$AT or $\dot{V}_{O_2}$Peak.
There was only a weak negative correlation between serum bilirubin and $\dot{V}_{O_2}$AT in our cohort.
These findings are similar to those of Parker and co-workers who reported that there was only a weak negative correlation between serum bilirubin and  $\dot{V}_{O_2}$Peak (Pearson's $\rho$ -0.21, p=0.02}) and no correlation with $\dot{V}_{O_2}$AT ($\rho$ = -0.15, p=NS) in patients undergoing pancreaticoduodenectomy \parencite{parker_serum_2014}. 
Junejo and co-workers studied oxygen extraction in 9 jaundiced patients during cardiopulmonary exercise test by placing catheters within the femoral vein. 
They reported that oxygen extraction was normal during cardiopulmonary exercise even in the presence of jaundice \parencite{junejo_peripheral_2014}

Taken together, these results appear to suggest that preoperative biliary drainage and resolution of jaundice are unlikely to improve aerobic capacity and cardiopulmonary physiology in these patients.

%Obstructive jaundice and systemic inflammation
However, jaundice was associated with elevated preoperative systemic inflammation as measured by C-reactive protein levels and serum albumin. 
Obstructive jaundice is known to be associated with wide ranging abnormalities of both humoral and innate immune systems \parencite{nehez_compromise_2002}. 
Both cellular immunity and humoral immunity are adversely affected by obstructive jaundice \parencite{padillo_cytokines_2001, scott-conner_pathophysiology_1994}. 
It is unclear if this immune dysfunction before surgery is associated with an abnormal postoperative systemic inflammatory response and this is explored further in Chapter \ref{ch_pre_post_sirs}.
However, the immunological abnormalities associated with obstructive jaundice do not appear to improve after biliary drainage \parencite{kimmings_endotoxin_2000}.

Elevated preoperative CRP was also independently associated with a low $\dot{V}_{O_2}$AT suggesting that impaired aerobic capacity may be associated with a systemic inflammatory response.
This finding is similar to that reported by Sultan and co-workers who found that low $\dot{V}_{O_2}$AT was independently associated with an elevated preoperative neutrophil-lymphocyte ratio in patients undergoing colorectal surgery \parencite{sultan_cardiopulmonary_2014}.

%CPET vs Hemoglobin
The association between low haemoglobin and low $\dot{V}_{O_2}$Peak would appear to be more easily explained since the oxygen carrying capacity of blood depends on haemoglobin concentration.  
Indeed, the linear relationship between oxygen consumption during exercise and haemoglobin concentration has been shown both in disease states \parencite{agostoni_relationship_2010} as well as in healthy volunteers \parencite{dellweg_cardiopulmonary_2008}.
This would suggest that maintaining optimal haemoglobin levels in the perioperative period will improve oxygen delivery and perhaps outcome as well. 
Excessive intra-operative blood loss has been reported to be associated with postoperative complications \parencite{pratt_risk_2008} and this may in part be due impaired oxygen delivery at a time of maximal physiological burden.

%CPET vs BMI, Sex
In the present study, it was of interest that $\dot{V}_{O_2}$, both at anaerobic threshold and at peak exercise, was significantly lower in females as well as in patients with  a high body mass index.
The basis of this relationship is not clear and may be related to body composition differences between these groups.
However, such an association has been previously reported \parencite{horwich_relationship_2009}.
This may reflect the difficulty in obtaining accurate $\dot{V}_{O_2}$ values in obese patients as a result of the calculations involved rather than due to true cardiopulmonary dysfunction. 
Other authors have suggested that different thresholds for CPET parameters may have to be considered in obese patients to improve risk-prediction \parencite{donnelly_criteria_1990,hulens_exercise_2001}.
Presence of cancer was also an independent predictor of low $\dot{V}_{O_2}$AT.
Sarcopenia is associated with reduced aerobic capacity \parencite{evans_sarcopenia_1993} and is common in pancreatic cancer \parencite{joglekar_sarcopenia_2015}.
This is the subject of further study in Chapter \ref{ch_bodycomp}.

%\subsection{Conclusion}
Obstructive jaundice, including severe obstructive jaundice did not affect preoperative cardiopulmonary exercise physiology. 
Therefore, preoperative biliary drainage cannot be expected to improve aerobic capacity in patients undergoing pancreaticoduodenectomy.
Both obstructive jaundice and low $\dot{V}_{O_2}$AT were associated with elevated preoperative systemic inflammation and provide further evidence of the role of jaundice and impaired aerobic capacity in immune dysregulation in these patients.
High BMI and female sex were independently associated with low $\dot{V}_{O_2}$AT and $\dot{V}_{O_2}$Peak and these parameters must be interpreted with caution in these subgroups.