### 2015-08-22 Tables/c1_cpet_parameters.tex

\begin{table}[p]
	\centering
	\caption{Parameters measured at spirometry.}
	\label{table:spirometry}
	\renewcommand{\arraystretch}{1.5} %Increases space between rows
	\setlength{\tabcolsep}{10pt} %sets the space between columns
	\begin{tabular}{| m{2cm}  m{2cm}  m{7.7cm} |}
		\hline
		Parameter & Units  & Description                          \\ \hline
		FVC       & litres & Forced Vital Capacity                \\
		FEV1      & litres & Forced Expiratory Volume in 1 second \\
		FEV1/FVC  & \%     & Tiffeneau-Pinelli index              \\ \hline
	\end{tabular}

	\vspace{2cm}

	\caption{Common parameters measured at cardiopulmonary exercise testing.}
	\label{table:cpet_parameters}
	\begin{tabular}{| m{2cm}  m{2cm}  m{7.7cm} |}
		\hline
		Parameter                  & Units      & Description                         \\ \hline
		Load                       & Watts      & Exercise Workload                   \\
		$\dot{V}_E$                & litres/min & Ventilatory Equivalent              \\
		Vt                         & litres     & Tidal volume                        \\
		$\dot{V}_{O_2}$            & litres/min & Absolute Oxygen uptake/consumption  \\
		$\dot{V}_{O_2}$/kg         & ml//kg/min & Corrected Oxygen uptake/consumption \\
		$\dot{V}_E/\dot{V}_{O_2}$  &            & Ventilatory Equivalent for $O_2$    \\
		$\dot{V}_{CO_2}$           & litres/min & Carbon-dioxide output               \\
		$\dot{V}_E/\dot{V}_{CO_2}$ &            & Ventilatory Equivalent for $CO_2$   \\
		RER                        &            & Respiratory Exchange Ratio          \\
		$PET_{O_2}$                & mmHg       & End Tidal $O_2$                     \\
		$PET_{CO_2}$               & mmHg       & End Tidal $CO_2$                    \\
		${O_2}$Pulse               & ml/beat    & Oxygen pulse                        \\
		HR                         & beats/min  & Heart Rate                          \\
		$B_f$                      & resps/min  & Breathing Frequency                 \\ \hline
	\end{tabular}

\end{table}