### 2015-08-22 Tables/c1_isgps.tex

\begin{sidewaystable}[htbp]
\centering
\footnotesize
\caption{Postoperative pancreatic fistula: ISGPF definition.}
\label{table:isgps_popf}
\renewcommand{\arraystretch}{1.2} %Increases space between rows
%\setlength{\tabcolsep}{10pt} %sets the space between columns
\begin{tabular}{|p{4.5cm}|p{4cm}|p{5cm}|p{5cm}|}
	\hline
	Grade                                & A        & B                 & C                                                   \\ \hline
	Clinical conditions                  & Well     & Often well        & Ill appearing/bad                                   \\
	Specific treatment                   & No       & Yes/no            & Yes                                                 \\
	US/CT (if obtained)                  & Negative & Negative/positive & Positive                                            \\
	Persistent drainage (after 3 weeks)† & No       & Usually yes       & Yes                                                 \\
	Reoperation                          & No       & No                & Yes                                                 \\
	Death related to POPF                & No       & No                & Possibly yes                                        \\
	Signs of infections                  & No       & Yes               & Yes                                                 \\
	Sepsis                               & No       & No                & Yes                                                 \\
	Readmission                          & No       & Yes/no            & Yes/no                                              \\ \hline
\end{tabular}

\vspace{0.7cm}

\caption{Postpancreatectomy haemorrhage: ISGPS definition.}
\label{table:isgps_pph}
\begin{tabular}{|m{4.5cm}|m{4cm}|m{5cm}|m{5cm}|}
	\hline
	Grade                                                             & A                                                                                & B                                                                                                                             & C                                                                                         \\ \hline
	Time of onset, location, severity and clinical impact of bleeding & Early, intra- or extraluminal, mild                                              & Early, intra- or extraluminal, severe	or Late, intra- or extraluminal, mild                                                   & Late, intra- or extraluminal, severe                                                      \\
	Clinical condition                                                & Well                                                                             & Often well/ intermediate, very rarely life-threatening                                                                        & Life-threatening, severely impaired                                                      \\
	Diagnostic consequence                                            & Observation, blood count, ultrasonography and, if necessary, computed tomography & Observation, blood count, ultrasonography, endoscopy, computed tomography, angiography                                      & Angiography, endoscopy, computed tomography                                              \\
	Therapeutic consequence                                           & No                                                                               & Transfusion of fluid/blood, intermediate care unit (or ICU), therapeutic endoscopy,† embolization, relaparotomy for early PPH & Localization of bleeding, angiography and embolization, (endoscopy†) or relaparotomy, ICU \\ \hline
\end{tabular}
\end{sidewaystable}