### 2015-08-23 Appendices/AppendixAccessDatabase.tex
Review - Minor changes to Appendices
- Modify short title
- insert \clearpage

\chapter{Visual Basic for Applications For Data Collection}
\label{AppendixAccessDatabase}
\lhead{Appendix \ref{AppendixAccessDatabase}. \emph{MS Access VBA For Data Collection}}
\clearpage
%------------------------------------------

\setstretch{1.0}