### 2015-08-23 Appendices/AppendixCPETRawData.tex

\chapter{Breath-by-breath CPET sample data}
\label{AppendixCPETRawData}
\lhead{Appendix \ref{AppendixCPETRawData}. \emph{Raw CPET Data}}
\clearpage
%------------------------------------------

\begin{figure}
	\centering
	\caption{Breath-by-breath sample data with values averaged every 10 seconds - Part 1.}
	\includegraphics[width=\textwidth,page=1]{Figures/cpet_9p_rawdata.pdf}
\end{figure}
\begin{figure}
	\centering
	\caption{Breath-by-breath sample data with values averaged every 10 seconds - Part 2.}
	\includegraphics[width=\textwidth,page=2]{Figures/cpet_9p_rawdata.pdf}
\end{figure}