### 2015-08-25 Chapters/Discussion.tex
Discussion - More updates
- See text
- This is a partial updata
- Proabably 2 or 3 more updates

% Discussion

\chapter{Discussion}
\label{ch_discussion}

\lhead{Chapter \ref{ch_discussion}. \emph{Discussion}} % This is for the header on each page - perhaps a shortened title

\clearpage
%----------------------------------------------------------------------------------------
\section{Aims of thesis}

The overall aim of the thesis was to examine the inter-relationships between preoperative clinico-pathological characteristics including cardiopulmonary exercise physiology, obstructive jaundice, body composition and preoperative systemic inflammation and post-operative complications and the post-surgical systemic inflammatory response in patients undergoing pancreaticoduodenectomy.

The present work also examined the factors that affect cardiopulmonary exercise physiology in order to enable a better understanding of the complex pathophysiology in these patients that is a consequence of the interaction of the patient's chronic conditions with the acute derangements brought on by pancreatic disease, including obstructive jaundice.

%################################################################################
\section{Clinical utility of CPET in predicting postoperative complications}
%################################################################################

\subsection{Risk assessment and predicting complications}

The results presented in chapter \ref{ch_cpet_outcomes} demonstrated that cardiopulmonary exercise testing has a role in identifying high risk patients who are more likely to develop complications, stay longer in hospital and are less likely to receive adjuvant therapy after pancreaticoduodenectomy. 
These findings have since been replicated by other authors examining the role of cardiopulmonary exercise testing in pancreatic surgery.

Ausania and co-workers evaluated the role of cardiopulmonary exercise testing in 122 patients who underwent a pancreaticoduodenectomy \parencite{ausania_effects_2012}
Low $\dot{V}_{O_2}$AT ($<$10.1 ml/kg/min) was the only independent predictor of a postoperative pancreatic fistula.
The incidence of a postoperative pancreatic fistula was 45\% in patients with a low $\dot{V}_{O_2}$AT while it was only 19\% in patients with a normal $\dot{V}_{O_2}$AT (p=0.02).
The association between $\dot{V}_{O_2}$AT and postoperative pancreatic fistula was independent of pancreatic duct size, body mass index, obstructive jaundice or preoperative biliary drainage.
They also reported in a separate study of 50 patients that complication  rates were higher in patients with a low $\dot{V}_{O_2}$AT undergoing palliative surgical bypass for advanced pancreatic cancer \parencite{ausania_double_2012}.

Junejo and co-workers identified elevated $\dot{V}_E/\dot{V}_{CO_2}$ to be an independent predictor of 30-day mortality after pancreaticoduodenectomy in their study that included 64 patients who had undergone cardiopulmonary exercise testing \parencite{junejo_cardiopulmonary_2014}.
$\dot{V}_{O_2}$AT and $\dot{V}_{O_2}$Peak  were not related to mortality or complications in this study.
However, they also noted that a high $\dot{V}_E/\dot{V}_{CO_2}$ was also associated with poor long-term survival.

\subsection{Assessing impact of prehabilitation or neoadjuvant treatment}
The results of our work as well as that of several others over the past 2 decades support the use of cardiopulmonary exercise testing as a clinical risk assessment tool in patients undergoing major surgery.
However, as presented in the subsequent chapters of this thesis, a better understanding of the determinants of aerobic capacity as well as the perioperative systemic inflammatory response will enable clinicians to identify high risk patients and optimise their perioperative care.

In a recent study of 8266 patients undergoing pancreaticoduodenectomy in the United States, Tzeng and co-workers reported that a third (3033) were of borderline fitness as a consequence of advanced age ($>$ 80), poor performance status, weight loss $>$ 10\%, pulmonary disease, recent myocardial infarction/angina, stroke history, and/or preoperative sepsis. 
Major complications (31.3 vs. 26.2\%) and mortality (4.1 vs. 2.3\%) were greater in patients with borderline fitness \parencite{tzeng_morbidity_2014}.
The authors recommended that surgeons identify these patients early, institute interventions to optimise their comorbidities and enrol these patients for prehabilitation.

Neoadjuvant treatments were associated with significant reduction in $\dot{V}_{O_2}$AT and $\dot{V}_{O_2}$Peak in patients with rectal cancer \parencite{west_effects_2014} and oesphago-gastric cancer \parencite{jack_effect_2014}.
West and co-workers used prehabilitation to return aerobic capacity to baseline levels in patients with locally advanced rectal cancer undergoing neo-adjuvant chemoradiotherapy.
They used cardiopulmonary exercise testing to monitor the impact of neo-adjuvant treatment and prehabilitation \parencite{west_effect_2015}.
Neoadjuvant therapy in borderline resectable pancreatic cancer is the subject of several trials with generally promising results \parencite{evans_preoperative_2008, gillen_preoperative/neoadjuvant_2010, golcher_neoadjuvant_2015, kim_multi-institutional_2013, sahora_phase_2014}.

It is likely that the impact of such treatment on fitness in these patients will be similar to that observed in patients with colorectal and oesophago-gastric cancer.
Prehabilitation would play an important role in improving outcomes in these patients while cardiopulmonary exercise testing will serve as an important clinical tool to objectively measure the impact of prehabilitation.

\subsection{Assessing impact on long-term survival}

Poor aerobic capacity is associated with reduced long-term survival after abdominal aortic aneurysm surgery \parencite{tang_cardiopulmonary_2012, grant_cardiopulmonary_2015}.
There is some evidence that cardiopulmonary exercise testing may predict long-term survival after pancreaticoduodenectomy for pancreatic cancer \parencite{junejo_cardiopulmonary_2014}.
Reduced aerobic capacity was also associated with poor survival in patients with lung cancer \parencite{jones_peak_2010}.

While the an elevated preoperative systemic inflammatory response has been shown to have an adverse impact on long-term survival in patients with pancreatic cancer both in those who underwent potentially curative surgery or palliation, the impact of aerobic capacity and comorbidity on long-term survival in patients with pancreatic cancer is less clear.
Some have suggested that comorbidity was associated with shorter survival in patients with pancreatic cancer \parencite{dias-santos_charlson_2015}.
However, other authors have reported that the relatively short median survival after pancreatic cancer is not affected by comorbidity in patients who underwent potentially curative surgery \parencite{kos_evaluation_2014}

The relationship between aerobic capacity and cancer-survival is important as it provides for a relatively easy therapeutic target.

% Different surgeries and patient groups may need different thresholds for risk prediction
% Done - CPET and long-term survival in PDAC
% Why is low $\dot{V}_{O_2}$ likely to leak - pancreatic perfusion 
%high lactate - popf de_schryver_early_2015

%karoliska - ansorge_early_2012
- better understanding of physiology
- role of sytemic inflammation \parencite{van_heek_hospital_2005, ho_effect_2003, birkmeyer_surgeon_2003, halm_is_2002}
- impact of sarcopenic-obesity \parencite{joglekar_sarcopenia_2015, reisinger_sarcopenia_2015, gonzalez_obesity_2014}
- role of tumour related factors \parencite{williams_surgical_2014}

%################################################################################
\section{Prehabilitation,improving aerobic capacity and mitigating risk}
%################################################################################
In Chapters \ref{ch_cpet_jaundice} and \ref{ch_bodycomp} we examined the preoperative patient factors that had an adverse effect on aerobic capacity.
A better understanding of the factors that affect $\dot{V}_{O_2}$ may identify potential therapeutic targets or modifiable risk factors that may be used to improve patient fitness and mitigate the risks of surgery.
We found that female sex, high body mass index, anaemia, presence of cancer and elevated CRP were all associated with either a low $\dot{V}_{O_2}$AT or a low $\dot{V}_{O_2}$Peak.

%################################################################################
\subsection{Obstructive jaundice}
%################################################################################
In our opinion, the most important observation made in chapter \ref{ch_cpet_jaundice} is the lack of association between obstructive jaundice and poor aerobic fitness. 
In chapter \ref{ch_outcomes} we reported that obstructive jaundice was not associated with increased incidence of postoperative complications or prolonged hospital stay.

We are aware of only one other published study looking at cardiopulmonary exercise testing and peripheral oxygen extraction in jaundiced versus non-jaundiced patients undergoing pancreaticoduodenectomy. 
Junejo and co-workers reported that peripheral oxygen extraction was normal during exercise in patients with malignant obstructive jaundice.\parencite{junejo_peripheral_2014} 
Preoperative biliary drainage is unlikely to improve aerobic fitness or modify their cardiopulmonary response to exercise/surgery \parencite{parker_serum_2014}.

We also demonstrated the immunosuppressive effects of obstructive jaundice on the postoperative systemic inflammatory response in chapter \ref{ch_pre_post_sirs}. 
To our knowledge, this is the first detailed study of the postoperative systemic inflammatory response in jaundiced versus non-jaundiced patients undergoing pancreaticoduodenectomy.

In their, appropriately titled paper \textit{`The Quandary of Preresection Biliary Drainage for Pancreatic Cancer'}, the authors discuss the complex physiological effects of obstructive jaundice, its impact on immune function and the differences between findings in animal studies and humans \parencite{tol_quandary_2012}.
However, they also review recent literature where there appears to be overwhelming evidence that preoperative biliary drainage increases the incidence of postoperative complications, especially infective complications. \parencite{van_der_gaag_preoperative_2010, arkadopoulos_preoperative_2014, fujii_preoperative_2015, furukawa_negative_2015}.

Our results therefore support the emerging view that routine preoperative biliary drainage should no longer be performed in all patients scheduled to undergo a pancreaticoduodenectomy.

%################################################################################
\subsection{Systemic inflammation}
%################################################################################
We reported on the preoperative factors affecting postoperative systemic inflammation in chapter \ref{ch_pre_post_sirs} and the clinical utility of monitoring trends in the early postoperative systemic inflammatory response in predicting complications in chapter \ref{ch_crp_comp}.

Preoperative systemic inflammation is increasingly recognised as playing an important role in determining not only long-term outcomes but also short-term outcomes after cancer surgery \parencite{kubo_elevated_2013, mohri_correlation_2014, moyes_preoperative_2009, neal_preoperative_2011, vashist_glasgow_2010,  haupt_association_1997}.
While we did not find an association between elevated preoperative systemic inflammation in our cohort of patients, we demonstrated in chapter \ref{ch_pre_post_sirs} that preoperative systemic inflammation had a significant impact on  the magnitude of the postoperative inflammatory response.
This supports the theory that abnormal systemic inflammation is a continuum that starts in the preoperative phase and continues into the postoperative phase.

It would appear that preoperative patient factors affect the postoperative systemic inflammatory response independent of intra-operative factors or postoperative complications. 
The mechanisms underlying the association between pre- and post-operative systemic inflammation require further study.

Several recent studies have demonstrated the value of monitoring trends in postoperative CRP levels in predicting complications after pancreaticoduodenectomy.
These studies performed at large, specialist centres, have used a combination of postoperative CRP and drain amylase levels to identify patients who are at the least risk of developing a Grade B/C pancreatic fistula \parencite{hiyoshi_usefulness_2013, ansorge_diagnostic_2014, kosaka_multivariate_2014}.

Our approach to interpreting the postoperative systemic inflammatory reponse has been different from the above mentioned studies. 
We demonstrated in chapter \ref{ch_crp_comp} that postoperative CRP did not predict the severity of a pancreatic fistula and did not have any clinical utility in the prediction of a POPF (section \ref{Postoperative CRP vs. POPF}).
However, we demonstrated a clear role for the use of postoperative CRP in predicting infective complications in the absence of a POPF.

In summary the combination of a low CRP and absence of POPF reduces the likelihood of any other infective complications and identifies patients who are suitable for early discharge.
Such patients will also be more likely to benefit from a full course of adjuvant chemotherapy although this requires further study.

Postoperative complications and the associated systemic inflammatory response are associated with early recurrence and poor-survival after colorectal \parencite{artinyan_infectious_2015, mcardle_impact_2005}, gastric \parencite{hayashi_impact_2015,kubota_prognostic_2014}  and pancreatic cancer surgery \parencite{aoyama_impact_2015, kamphues_postoperative_2012}.

Early recognitioon and aggressive management of complications is therefore important in mitigating the harmful long-term effects of such adverse events and the associated inflammation.

%################################################################################
\subsection{Body composition}
%################################################################################
Overweight and obese patients had significantly lower $\dot{V}_{O_2}$AT and $\dot{V}_{O_2}$Peak in spite of no significant increase in known cardiac or respiratory comorbidities. 
We present a detailed analysis of the relationship between body composition and preoperative cardiopulmonary exercise testing in chapter \ref{ch_bodycomp}.
This has not been reported before in surgical patients to the best of out knowledge.

The loss of skeletal muscle mass in patients with pancreatic cancer and spurious correlation with obesity due to correction for total body weight may partly explain the low $\dot{V}_{O_2}$ in these patients. 
Low aerobic capacity as measured by cardiopulmonary exercise testing must be interpreted with caution in overweight/obese patients and in patients with pancreatic cancer. 
This is especially true for $\dot{V}_{O_2}$ that has been scaled for total body weight.
Further studies must evaluate other parameters such as $O_2$Pulse or $\dot{V}_E/\dot{V}_{O_2}$ for their ability to predict postoperative outcomes. 
With the significant increase in the proportion of obese patients undergoing major surgery and the presence of sarcopenic-obesity in patients with pancreatic cancer, poor performance at cardiopulmonary exercise testing may not necessarily be related to cardiac or pulmonary function.

%################################################################################
\subsection{Aerobic exercise}
%################################################################################

West and co-workers used prehabilitation to return aerobic capacity to baseline levels in patients with locally advanced rectal cancer undergoing neo-adjuvant chemoradiotherapy \parencite{west_effect_2015}.
Moderate aerobic and resistance exercises and protein supplementation started 4 weeks before surgery was noted to improve postoperative functional exercise capacity in patients undergoing colorectal cancer surgery \parencite{gillis_prehabilitation_2014}.
Exercise programs as short as 4 weeks in duration have been reported to improve aerobic capacity by upto 10\%  measured objectively by cardiopulmonary exercise testing.
This could potentially move approximately 30\% of high risk patients into the low risk category \parencite{dunne_pmo-029_2012}.

A recent consensus statement by the `Exercise Prehabilitation in Colorectal Cancer Delphi Study Group' in the United Kingdom showed agreement between colorectal surgeons that prehabilitation must form an important part of the preparation of elective patients for major surgery \parencite{boereboom_forming_2015}. 
The surgeons also agreed that improving aerobic capacity is likely to improve outcomes but recommended that further studies were needed before this became established clinical practice.

%################################################################################
\subsection{Anaemia}
%################################################################################

Measures other than exercise may also improve aerobic capacity. 
Correction of anaemia was associated with an improvement in $\dot{V}_{O_2}$AT by 0.39 ml/kg/min for every 1 gm/dl improvement in haemoglobin \parencite{wright_cardiopulmonary_2014}. 
While anaemia did not change the effect of aerobic exercise in improving physical performance, the overall aerobic fitness was significantly lower in anaemic patients than in patients with normal haemoglobin \parencite{bellotto_anemia_2011}.
However, perioperative blood transfusion in pancreatic cancer patients was associated with poorer long-term outcomes \parencite{kneuertz_effects_2011, sutton_perioperative_2014}.

While iron supplementation is associated with correction of anaemia and reduction in the requirement for blood transfusion, the role of preoperative iron supplementation in improving outcomes is unclear. 
Further randomised controlled trials investigating the short-term and long-term consequences of preoperative iron supplementation in patients undergoing cancer surgery are required before this becomes accepted clinical practice \parencite{beris_perioperative_2008,hallet_impact_2014}.

%################################################################################
Moreover, low aerobic capacity was not associated with documented cardiovascular or respiratory comorbidity or with an elevated POSSUM physiology score.
It is possible that most patients with significant cardiac or respiratory comorbidity were identified as unsuitable for surgery and therefore were never subject to cardiopulmonary exercise testing. 
Alternatively, it is possible that these patients did very poorly at cardiopulmonary exercise testing and this, in addition to their cardiorespiratory medical history precluded them from surgery.
We did not include any patients who did not have an operation and this may have introduced a selection bias.
%################################################################################

%################################################################################
\section{Improving outcomes after pancreaticoduodenectomy}
%################################################################################

\subsection{Centralisation and Specialisation}

\subsection{Early and aggressive management of complications}

\subsection{Enhanced Recovery Programmes}

- centralisation
- specialisation
- early rescue and aggressive management of complications \parencite{gouma_rates_2000}

\subsection{Prehabilitation}

%using cpet to assess impact of neo-adjuvant treatment
%recommend prehab for neoadj in pancreatic cancer
%different types of surgery may need different levels of preparation

%boereboom_forming_2015

\subsection{Intra-operative care}

The results of chapter \ref{ch_cpet_outcomes} and the work of Ausania and co-workers appeared to suggest a link between reduced oxygen delivery and increased incidence of pancreatic fistula. 
We hypothesised that tissue hypoperfusion may have played a role in the poor healing of the pancreatico-jejunal anastomosis.

Reyad and co-workers performed a randomised trial to examine the effect of intra-operative dobutamine infusion during pancreaticoduodenectomy on splanchnic perfusion, hemodynamics, and overall postoperative outcome \parencite{reyad_effect_2013}. 
They reported that intra-operative dobutamine use was associated with improved global oxygen delivery as measured by arterial and venous blood gases, splanchnic perfusion as measured by gastric tonometry and postoperative complications.
It was interesting to note that the incidence of anastomotic leak was 30\% in the control group, 15\% in the group that received 3$\mu$g/kg/min of dobutamine and 5\% in the group that received 5$\mu$g/kg/min of dobutamine (p$<$0.05). The overall complication rate also decreased from 70\% in the control group to 40\% and 20\% in the dobutamine groups (p$<0.05$).

However, the benefit of goal-directed therapy to improve organ perfusion and oxygen delivery in improving postoperative outcomes remains unclear \parencite{grocott_perioperative_2013, pearse_effect_2014}. 
A recent randomised controlled trial evaluated the role of goal-directed therapy in high-risk patients undergoing major elective surgery that targeted individualised oxygen delivery \parencite{ackland_individualised_2015}.
The study did show that patients whose postoperative oxygen delivery was similar to their preoperative values had fewer complications.
However, oxygen delivery was not influenced by goal-directed therapy as any beneficial effect of the intervention was lost with the autonomic nervous system changes that accompanied the increased intravenous fluids and inotropes in the treatment cohort.

\subsection{Post-operative care}

\section{Future directions}

%Early recognition of patients at increased risk of postoperative complications facilitates prompt initiation of appropriate treatment and judicious allocation of critical care resources. 
%Moreover, identification of patients who are unlikely to develop a major postoperative complication allows their progress on enhanced recovery pathways with the benefits of rapid recovery, reduced morbidity and early discharge. 
%These patients are also more likely to receive adjuvant therapy without undue delay. 
%Recognition of these low risk patients helps improve outcomes, reduces hospital costs and releases valuable resources for the care of high risk patients. 
%The results reported in Chapter [ss] demonstrate the value of early post-operative C-reactive protein in identifying patients who are more likely to develop an infective complication. 
%The results show that CRP is only useful in the absence of a postoperative pancreatic fistula. 
%The combination of low drain amylase and low CRP on the third postoperative day identifies patients who do not have a pancreatic fistula and are less likely to develop an infective complication.