### 2015-08-25 Chapters/Discussion.tex
Discussion - Complete
- Final big commit
- Next few commits will all be just minor fixes and formatting

% Discussion

\chapter{Discussion}
\label{ch_discussion}

\lhead{Chapter \ref{ch_discussion}. \emph{Discussion}} % This is for the header on each page - perhaps a shortened title

\clearpage
%----------------------------------------------------------------------------------------
\section{Summary of Aims}

The overall aim of the thesis was to examine the inter-relationships between preoperative clinico-pathological characteristics including cardiopulmonary exercise physiology, obstructive jaundice, body composition and preoperative systemic inflammation and post-operative complications and the post-surgical systemic inflammatory response in patients undergoing pancreaticoduodenectomy.

The present work also examined the factors that affect cardiopulmonary exercise physiology in order to enable a better understanding of the complex pathophysiology in these patients that is a consequence of the interaction of the patient's chronic conditions with the acute derangements brought on by pancreatic disease, including obstructive jaundice.

\clearpage

%################################################################################
\section{Clinical utility of CPET}
%################################################################################

\subsection{Risk assessment and predicting complications}

The results presented in chapter \ref{ch_cpet_outcomes} demonstrated that cardiopulmonary exercise testing has a role in identifying high risk patients who are more likely to develop complications, stay longer in hospital and are less likely to receive adjuvant therapy after pancreaticoduodenectomy. 
These findings have since been replicated by other authors examining the role of cardiopulmonary exercise testing in pancreatic surgery.

Ausania and co-workers evaluated the role of cardiopulmonary exercise testing in 122 patients who underwent a pancreaticoduodenectomy \parencite{ausania_effects_2012}
Low $\dot{V}_{O_2}$AT ($<$10.1 ml/kg/min) was the only independent predictor of a postoperative pancreatic fistula.
The incidence of a postoperative pancreatic fistula was 45\% in patients with a low $\dot{V}_{O_2}$AT while it was only 19\% in patients with a normal $\dot{V}_{O_2}$AT (p=0.02).
The association between $\dot{V}_{O_2}$AT and postoperative pancreatic fistula was independent of pancreatic duct size, body mass index, obstructive jaundice or preoperative biliary drainage.
They also reported in a separate study of 50 patients that complication  rates were higher in patients with a low $\dot{V}_{O_2}$AT undergoing palliative surgical bypass for advanced pancreatic cancer \parencite{ausania_double_2012}.

Junejo and co-workers identified elevated $\dot{V}_E/\dot{V}_{CO_2}$ to be an independent predictor of 30-day mortality after pancreaticoduodenectomy in their study that included 64 patients who had undergone cardiopulmonary exercise testing \parencite{junejo_cardiopulmonary_2014}.
%$\dot{V}_{O_2}$AT and $\dot{V}_{O_2}$Peak  were not related to mortality or complications in this study.
However, they also noted that a high $\dot{V}_E/\dot{V}_{CO_2}$ was also associated with poor long-term survival.

The results of our work in chapter \ref{ch_cpet_outcomes} as well as that of several others over the past 2 decades support the use of cardiopulmonary exercise testing as a clinical risk assessment tool in patients undergoing major surgery.

\subsection{Assessing impact of prehabilitation and neoadjuvant treatment}

Our results also suggest that a better understanding of the determinants of aerobic capacity as well as the perioperative systemic inflammatory response will enable clinicians to identify high risk patients and optimise their perioperative care.

In a recent study of 8266 patients undergoing pancreaticoduodenectomy in the United States, Tzeng and co-workers reported that a third (3033) were of borderline fitness as a consequence of advanced age ($>$ 80 years), poor performance status, weight loss $>$ 10\%, pulmonary disease, recent myocardial infarction/angina, stroke history, and/or preoperative sepsis. 
Major complications (31.3 vs. 26.2\%) and mortality (4.1 vs. 2.3\%) were greater in patients with borderline fitness \parencite{tzeng_morbidity_2014}.
The authors recommended that surgeons identify these patients early, institute interventions to optimise their comorbidities and enrol these patients for prehabilitation.

Neoadjuvant treatments were associated with significant reduction in $\dot{V}_{O_2}$AT and $\dot{V}_{O_2}$Peak in patients with rectal cancer \parencite{west_effects_2014} and oesphago-gastric cancer \parencite{jack_effect_2014}.
West and co-workers used prehabilitation to return aerobic capacity, as measured by cardiopulmonary exercise testing, to baseline levels in patients with locally advanced rectal cancer undergoing neo-adjuvant chemoradiotherapy \parencite{west_effect_2015}.
Neoadjuvant therapy in borderline resectable pancreatic cancer is the subject of several trials with generally promising results \parencite{evans_preoperative_2008, gillen_preoperative/neoadjuvant_2010, golcher_neoadjuvant_2015, kim_multi-institutional_2013, sahora_phase_2014}.

The impact of neoadjuvant treatment on fitness in patients with pancreatic cancer are likely to be similar or worse to that observed in other cancer patients.
Prehabilitation can play an important role in improving outcomes in these patients while cardiopulmonary exercise testing serves as an objective clinical tool to quantify the impact of prehabilitation.

\subsection{Assessing impact on long-term survival}

There is some evidence that cardiopulmonary exercise testing may predict long-term survival after pancreaticoduodenectomy for pancreatic cancer \parencite{junejo_cardiopulmonary_2014}.
Reduced aerobic capacity was also associated with poor survival in patients with lung cancer \parencite{jones_peak_2010} and reduced long-term survival after abdominal aortic aneurysm surgery \parencite{tang_cardiopulmonary_2012, grant_cardiopulmonary_2015}.

While an elevated preoperative systemic inflammatory response was associated with shorter survival in patients with pancreatic cancer, both after potentially curative surgery \parencite{jamieson_systemic_2005, torre_glasgow_2012} or palliation for advanced disease \parencite{glen_evaluation_2006}, the impact of aerobic capacity on long-term survival in patients with pancreatic cancer is less clear.
Some have suggested that comorbidity was associated with shorter survival in patients with pancreatic cancer \parencite{dias-santos_charlson_2015}.
However, other authors have reported that the relatively short median survival after pancreatic cancer surgery is not affected by comorbidity whose effects are more long-term \parencite{kos_evaluation_2014}

Further studies examining the relationship between aerobic capacity measured using objective methods and long-term survival are important. 
Aerobic capacity is a relatively easy therapeutic target that may have the potential to improve not only short-term but also long-term outcomes.

%################################################################################
\section{Prehabilitation, improving aerobic capacity and mitigating risk}
\label{disc_prehabilitation}
%################################################################################
In Chapters \ref{ch_cpet_jaundice} and \ref{ch_bodycomp} we examined the preoperative patient factors that had an adverse effect on aerobic capacity.
A better understanding of the factors that affect $\dot{V}_{O_2}$ may identify potential therapeutic targets or modifiable risk factors that may be used to improve patient fitness and mitigate the risks of surgery.
We found that female sex, high body mass index, anaemia, presence of cancer and elevated CRP were all associated with either a low $\dot{V}_{O_2}$AT or a low $\dot{V}_{O_2}$Peak.

%################################################################################
\subsection{Obstructive jaundice}
%################################################################################
In our opinion, the most important observation made in chapter \ref{ch_cpet_jaundice} was the lack of association between obstructive jaundice and poor cardiopulmonary function. 
We are aware of only one other similar study by Junejo and co-workers. They reported that peripheral oxygen extraction was normal during cardiopulmonary exercise in patients with malignant obstructive jaundice.\parencite{junejo_peripheral_2014} 
Preoperative biliary drainage is unlikely to improve aerobic fitness or modify their cardiopulmonary response to exercise/surgery \parencite{parker_serum_2014}.

We also demonstrated the immunosuppressive effects of obstructive jaundice on the postoperative systemic inflammatory response in chapter \ref{ch_pre_post_sirs}. 
To our knowledge, this is the first detailed study of the postoperative systemic inflammatory response in jaundiced versus non-jaundiced patients undergoing pancreaticoduodenectomy.

In their, appropriately titled paper \textit{`The Quandary of Preresection Biliary Drainage for Pancreatic Cancer'}, the authors discuss the complex physiological effects of obstructive jaundice, its impact on immune function and the differences between findings in animal studies and humans \parencite{tol_quandary_2012}.
However, they also review recent literature where there appears to be overwhelming evidence that preoperative biliary drainage increases the incidence of postoperative complications, especially infective complications. \parencite{van_der_gaag_preoperative_2010, arkadopoulos_preoperative_2014, fujii_preoperative_2015, furukawa_negative_2015}.

In chapter \ref{ch_outcomes} we reported that obstructive jaundice was not associated with increased incidence of postoperative complications or prolonged hospital stay. 
Taken together, our results support the emerging view that routine preoperative biliary drainage should no longer be performed in all patients scheduled to undergo a pancreaticoduodenectomy.

%################################################################################
\subsection{Systemic inflammation}
%################################################################################
We reported on the preoperative factors affecting postoperative systemic inflammation in chapter \ref{ch_pre_post_sirs} and the clinical utility of monitoring trends in the early postoperative systemic inflammatory response in predicting complications in chapter \ref{ch_crp_comp}.

Preoperative systemic inflammation is increasingly recognised as playing an important role in determining not only long-term outcomes but also short-term outcomes after cancer surgery \parencite{kubo_elevated_2013, mohri_correlation_2014, moyes_preoperative_2009, neal_preoperative_2011, vashist_glasgow_2010,  haupt_association_1997}.
While we did not find an association between elevated preoperative systemic inflammation and complications in our cohort of patients, we demonstrated in chapter \ref{ch_pre_post_sirs} that preoperative systemic inflammation had a significant impact on  the magnitude of the postoperative inflammatory response.
This supports the theory that abnormal systemic inflammation is a continuum that starts in the preoperative phase and continues into the postoperative phase.

It would appear that preoperative patient factors affect the postoperative systemic inflammatory response independent of intra-operative factors or postoperative complications. 
The mechanisms underlying the association between pre- and post-operative systemic inflammation require further study.

%################################################################################
\subsection{Body composition}
%################################################################################
Overweight and obese patients had significantly lower $\dot{V}_{O_2}$AT and $\dot{V}_{O_2}$Peak in spite of no significant increase in known cardiac or respiratory comorbidities. 
We present a detailed analysis of the relationship between body composition and preoperative cardiopulmonary exercise testing in chapter \ref{ch_bodycomp}.
This has not been reported before in surgical patients to the best of our knowledge.

The loss of skeletal muscle mass in patients with pancreatic cancer and spurious correlation with obesity due to correction for total body weight may partly explain the low $\dot{V}_{O_2}$ in these patients. 
Low aerobic capacity as measured by cardiopulmonary exercise testing must be interpreted with caution in overweight/obese patients and in patients with pancreatic cancer. 
This is especially true for $\dot{V}_{O_2}$ that has been scaled for total body weight.
Further studies must evaluate other parameters such as $O_2$Pulse or $\dot{V}_E/\dot{V}_{O_2}$ for their ability to predict postoperative outcomes. 
The proportion of obese patients undergoing major surgery is increasing. 
Moreover, sarcopenia and sarcopenic-obesity are recognised as an important adverse prognostic feature in pancreatic cancer \parencite{tan_sarcopenia_2009, peng_impact_2012} and is associated with increased postoperative complications after pancreaticoduodenectomy \parencite{joglekar_sarcopenia_2015, pausch_cachexia_2012}.
Sarcopenia was also an independent risk factor for an elevated postoperative systemic inflammatory response in patients with colorectal cancer \parencite{reisinger_sarcopenia_2015}.

Our results suggest that poor performance at cardiopulmonary exercise testing may not necessarily be related to only cardiac or pulmonary function. 
Body composition appears to not only affect aerobic capacity but also systemic inflammation.
Improvements in body composition in cancer patients are achievable through nutritional supplementation \parencite{machado_whey_2015} as well as increased physical activity.
Improvement in our understanding of cancer cachexia and the role of inflammation will also help optimise patients for treatment and prolong survival.

%################################################################################
\subsection{Anaemia}
%################################################################################

Measures other than exercise may also improve aerobic capacity. 
Correction of anaemia was associated with an improvement in $\dot{V}_{O_2}$AT by 0.39 ml/kg/min for every 1 gm/dl improvement in haemoglobin \parencite{wright_cardiopulmonary_2014}. 
While anaemia did not change the effect of aerobic exercise in improving physical performance, the overall aerobic fitness was significantly lower in anaemic patients than in patients with normal haemoglobin \parencite{bellotto_anemia_2011}.
However, perioperative blood transfusion in pancreatic cancer patients was associated with poorer long-term outcomes \parencite{kneuertz_effects_2011, sutton_perioperative_2014}.

While iron supplementation is associated with correction of anaemia and reduction in the requirement for blood transfusion, the role of preoperative iron supplementation in improving outcomes is unclear. 
Further randomised controlled trials investigating the short-term and long-term consequences of preoperative iron supplementation in patients undergoing cancer surgery are required before this becomes accepted clinical practice \parencite{beris_perioperative_2008,hallet_impact_2014}.

%################################################################################
\subsection{Aerobic exercise}
%################################################################################

West and co-workers used prehabilitation to return aerobic capacity to baseline levels in patients with locally advanced rectal cancer undergoing neo-adjuvant chemoradiotherapy \parencite{west_effect_2015}.
Moderate aerobic and resistance exercises and protein supplementation started 4 weeks before surgery was noted to improve postoperative functional exercise capacity in patients undergoing colorectal cancer surgery \parencite{gillis_prehabilitation_2014}.
Exercise programs as short as 4 weeks in duration have been reported to improve aerobic capacity by upto 10\%  measured objectively by cardiopulmonary exercise testing.
This could potentially move approximately 30\% of high risk patients into the low risk category \parencite{dunne_pmo-029_2012}.

A recent consensus statement by the `Exercise Prehabilitation in Colorectal Cancer Delphi Study Group' in the United Kingdom showed agreement between colorectal surgeons that prehabilitation must form an important part of the preparation of elective patients for major surgery \parencite{boereboom_forming_2015}. 
The surgeons also agreed that improving aerobic capacity is likely to improve outcomes but recommended that further studies were needed before this became established clinical practice.

%%################################################################################
%Moreover, low aerobic capacity was not associated with documented cardiovascular or respiratory comorbidity or with an elevated POSSUM physiology score.
%It is possible that most patients with significant cardiac or respiratory comorbidity were identified as unsuitable for surgery and therefore were never subject to cardiopulmonary exercise testing. 
%Alternatively, it is possible that these patients did very poorly at cardiopulmonary exercise testing and this, in addition to their cardiorespiratory medical history precluded them from surgery.
%We did not include any patients who did not have an operation and this may have introduced a selection bias.
%%################################################################################

%################################################################################
\section{Improving outcomes after pancreaticoduodenectomy}
%################################################################################

\subsection{Centralisation and Specialisation}

One of the biggest factors in improving both short-term and long-term outcomes after cancer surgery has been the centralisation of cancer treatment and the associated specialisation of surgeons.

Birkmeyer and co-workers were the first to report in 1999 that hospital volume was not only related to postoperative mortality and morbidity but also to long-term cancer survival in patients undergoing pancreaticoduodenectomy in the United States\parencite{birkmeyer_relationship_1999}.
Better overall quality of care, better cancer operations based on oncological principles, better rates of adjuvant therapy may have all contributed to better outcomes in high-volume centres. 
They also reported in 2003 that operative mortality was inversely related to surgeon volume emphasising the importance of surgical expertise in improving outcomes \parencite{birkmeyer_surgeon_2003}.
Gouma and co-workers reported very similar findings in the Netherlands were morbidity and mortality after pancreaticoduodenectomy were linked to hospital volume\parencite{gouma_rates_2000}.

The importance of multi-disciplinary teams in the administration of complex perioperative care in these patients by Ho and Heslin who reported that the overall duration with pancreaticoduodenectomy in a hospital was associated with better postoperative outcomes \parencite{ho_effect_2003}.
This finding stresses the importance of organisational processes and allied specialists including anaesthetists, intensivists, diagnostic/interventional radiologists, nurses, physiotherapists and dietitians.

However, in spite of these advances, significant variation in practice remains an obstacle in improving outcomes after pancreaticoduodenectomy for pancreatic cancer \parencite{cyr_canadian_2015}.

\subsection{Goal-directed therapy}

The results of chapter \ref{ch_cpet_outcomes} and the work of Ausania and co-workers \parencite{ausania_effects_2012} suggest a link between reduced oxygen delivery and increased incidence of pancreatic fistula. 
We hypothesised that tissue hypoperfusion may have contributed to poor healing of the pancreatico-jejunal anastomosis.

Reyad and co-workers performed a randomised trial of the effect of dobutamine infusion during pancreaticoduodenectomy on splanchnic perfusion, hemodynamics, and overall postoperative outcome \parencite{reyad_effect_2013}. 
They reported that intra-operative dobutamine use was associated with improved global oxygen delivery as measured by arterial and venous blood gases. 
Splanchnic perfusion as measured by gastric tonometry also improved.
It was interesting to note that the incidence of anastomotic leak was 30\% in the control group, 15\% in the group that received 3$\mu$g/kg/min of dobutamine and 5\% in the group that received 5$\mu$g/kg/min of dobutamine (p$<$0.05). 
The overall complication rate also decreased from 70\% in the control group to 40\% and 20\% in the dobutamine groups (p$<0.05$).

However, the benefit of goal-directed therapy to improve organ perfusion and oxygen delivery in improving postoperative outcomes remains unclear \parencite{grocott_perioperative_2013, pearse_effect_2014}. 
A recent randomised controlled trial evaluated the role of goal-directed therapy in high-risk patients undergoing major elective surgery that targeted individualised oxygen delivery \parencite{ackland_individualised_2015}.
The study did show that patients whose postoperative oxygen delivery was similar to their preoperative values had fewer complications.
However, oxygen delivery was not influenced by goal-directed therapy as any beneficial effect of the intervention was lost with the autonomic nervous system changes that accompanied the increased intravenous fluids and inotropes in the treatment cohort.

Improving oxygen delivery and organ perfusion in surgical patients remain the subjects of intense research.
However, the results presented in this thesis support the argument for therapeutic interventions aimed at improving oxygen delivery and organ perfusion before the patient with pancreatic cancer reaches the operating room as outlined in section \ref{disc_prehabilitation}.

\subsection{Early recognition and aggressive management of complications}

Postoperative complications and the associated systemic inflammatory response are associated with early recurrence and poor-survival after colorectal \parencite{artinyan_infectious_2015, mcardle_impact_2005}, gastric \parencite{hayashi_impact_2015,kubota_prognostic_2014}  and pancreatic cancer surgery \parencite{aoyama_impact_2015, kamphues_postoperative_2012}.

Early recognition and aggressive management of complications is therefore important in mitigating the harmful long-term effects of such adverse events and the associated inflammation. 
Several recent studies have demonstrated the value of monitoring trends in postoperative CRP levels in predicting complications after pancreaticoduodenectomy \parencite{hiyoshi_usefulness_2013, ansorge_diagnostic_2014, kosaka_multivariate_2014}.

Our approach to interpreting the postoperative systemic inflammatory response has been different from these studies. 
We demonstrated in chapter \ref{ch_crp_comp} that postoperative CRP did not predict the severity of a pancreatic fistula and did not have any clinical utility in the prediction of a POPF (section \ref{Postoperative CRP vs. POPF}).
However, we demonstrated a clear role for the use of postoperative CRP in predicting infective complications in the absence of a POPF. 
The combination of a low CRP and absence of POPF reduced the likelihood of any other infective complications and identified patients who may be suitable for early discharge.

`Failure to rescue' after postoperative complications is one of the factors responsible for the discrepancy between morbidity and mortality in low-volume versus high-volume centres.
Mortality rates in patients undergoing a variety of complex surgical procedures including pancreatectomy was 2.5 times higher in low-volume centres than high-volume centres. 
However, complication rates were similar between these hospitals. 
Failure to identify and manage complications aggressively may have resulted in the increased mortality in the low-volume hospitals \parencite{ghaferi_complications_2009}.
In a recent large study involving nearly 2 million surgical patients, Ferraris VA and co-workers studied the outcome of over two-hundred thousand patients who had complications. 
They noted that 20\% of high-risk patients accounted for 90\% of deaths due to postoperative complications. 
They also reported that a simple risk-scoring system based on preoperative characteristics was able to identify these patients. 
Early recognition of complications in these patients and aggressive management may result in improved outcomes in these high risk patients.
These findings are supported by other studies involving emergency abdominal surgery \parencite{sheetz_improving_2013} and patients at major trauma centres \parencite{haas_prevention_2011}

The results presented in this thesis not only allow identification of high risk patients before surgery but also help in the early identification of postoperative complications.

\section{Future directions}

The work done in this thesis raises a number of important clinical questions. 
Further studies on the inter-relationships between aerobic fitness, comorbidity, systemic inflammation at the metabolic level in patients with cancer will allow better understanding of their impact of postoperative outcomes. 
Objective measurement of aerobic fitness and systemic inflammation will allow for individualised perioperative care that starts at the time of diagnosis and continues in the postoperative period. 
Further studies will have to focus on improving modifiable risk factors such as $\dot{V}_{O_2}$ starting with community/home-based interventions and continuing with individualised goal-directed therapy during and after surgery.
Aggressive protocol-driven early rescue strategies that include identification of high risk patients before surgery, early recognition of complications and multi-disciplinary management may help improve postoperative outcomes and attenuate the adverse effects of complications on long-term survival.
This in combination with individualised cancer treatments both in the neo-adjuvant and adjuvant settings will help improve outcomes in patients with pancreatic cancer.