### 2015-08-27 Tables/crp_comp_CRP4_vs_LOS.tex

%12/07/15 - Started this table
\begin{table}[p]
	\centering
	\caption{The relationship between CRP on Day 4 and other postoperative events in patients with no POPF.}
	\label{table:crp_comp_CRP4_vs_LOS}
	\renewcommand{\arraystretch}{1.4} %Increases space between rows
	\setlength{\tabcolsep}{9pt} %sets the space between columns
	\begin{tabular}{|l l  c c r |}
		\hline
		                      &          & \multicolumn{2}{c}{Day 4 CRP (mg/l)} &  \\
		                      &          & $\leq$125   & $>$125                   & \textit{p}   \\ \hline
		Postoperative stay    & Days     & 12 (9 - 16) & 15 (13 - 26)             & $<$0.001$^a$ \\
		(days)                & $\leq$14 & 43 (66.2\%) & 30 (49.2\%)              & $<$0.054$^b$ \\
		                      & $>$14    & 22 (33.8\%) & 31 (50.8\%)              &  \\
		Critical care stay    & Days     & 6 (4 - 7)   & 7 (5 - 10)               & 0.001$^a$    \\
		(days)                & $\leq$7  & 57 (87.7\%) & 38 (62.3\%)              & 0.001$^b$    \\
		                      & $>$7     & 8 (12.3\%)  & 23 (37.7\%)              &  \\
		Crit. care admissions & 1        & 60 (92.3\%) & 48 (78.7\%)              & 0.029$^b$    \\
		                      & $>$1     & 5 (7.7\%)   & 13 (21.3\%)              &  \\
		Reoperation           & No       & 63 (96.9\%) & 57 (93.4\%)              & $<$0.359$^b$ \\
		                      & Yes      & 2 (3.1\%)   & 4 (6.6\%)                &  \\ \hline
		\multicolumn{5}{l}{a - Mann-Whitney U test; b - Chi-square test}
	\end{tabular}
\end{table}