### 2016-02-19 Tables/bc_at_new_indices.tex
\caption* in Body comp

\begin{sidewaystable}[p]
	\caption{The relationship between body composition, body habitus and $\dot{V}_{O_2}$AT scaled using different factors (controlled for sex) in patients undergoing major pancreatic surgery.}
	\label{table:bc_at_new_indices}
	\footnotesize
	\centering
	\renewcommand{\arraystretch}{1.5} %Increases space between rows
	%\setlength{\tabcolsep}{9pt} %sets the space between columns
	%\hline\noalign{\smallskip} %increases space after the line
	\begin{tabular}{|l|cc|cc|cc|cc|cc|cc|}
		\hline
		& \multicolumn{2}{c|}{Absolute $\dot{V}_{O_2}$} & \multicolumn{2}{c|}{$\frac{\dot{V}_{O_2}}{Weight}$} & \multicolumn{2}{c|}{$\frac{\dot{V}_{O_2}}{Height^2}$} & \multicolumn{2}{c|}{$\frac{\dot{V}_{O_2}}{BMI}$} & \multicolumn{2}{c|}{$\frac{\dot{V}_{O_2}}{Skeletal\ Muscle}$}  & \multicolumn{2}{c|}{$\frac{\dot{V}_{O_2}}{Lean\ Body\ Mass}$}\\

		Variable              	& $\rho$ &\textit{p}& $\rho$ &\textit{p}& $\rho$ &\textit{p}& $\rho$ &\textit{p}& $\rho$ &\textit{p}& $\rho$ &\textit{p} \\ \hline
		Subcut. fat (cm$^2$)     & 0.191  &  0.089   & -0.326 &  0.003   & 0.169  &  0.132   & -0.217 &  0.051   & 0.018  &  0.872   & -0.041 & 0.716 \\
		Visceral fat (cm$^2$)    & 0.184  &  0.103   & -0.365 &  0.001   & 0.154  &  0.168   & -0.279 &  0.012   & -0.065 &  0.563   & -0.108 & 0.339\\
		Total fat (cm$^2$)       & 0.231  &  0.038   & -0.400 & $<$0.001 & 0.191  &  0.088   & -0.286 &  0.010   & -0.021 &  0.851   & -0.082 & 0.466\\
		Skeletal Muscle (cm$^2$) & 0.463  & $<$0.001 & -0.078 &  0.487   & 0.329  &  0.003   & 0.105  &  0.353   & -0.429 & $<$0.001 & 0.133 & 0.237 \\
		Height (cm)               & 0.446  & $<$0.001 & 0.019  &  0.870   & 0.001  &  0.993   & 0.475  & $<$0.001 & 0.105  &  0.350   & 0.019 & 0.865\\
		Weight (kg)               & 0.545  & $<$0.001 & -0.302 &  0.006   & 0.358  &  0.001   & -0.043 &  0.703   & 0.001  &  0.995   & 0.033 & 0.770\\
		Body Mass Index (kg/m$^2$)      & 0.357  &  0.001   & -0.363 &  0.001   & 0.425  & $<$0.001 & -0.345 &  0.002   & -0.059 &  0.603   & 0.035 & 0.756\\ \hline
		\multicolumn{11}{l}{$\rho$ - Pearson's partial correlation adjusted for sex.}
	\end{tabular}
	\medskip
	\caption*{$\dot{V}_{O_2}$AT normalised using different measures of body habitus (including weight, square of height, body mass index, skeletal muscle area and estimated lean body mass) was compared to body composition as well as weight, height and BMI in an exploratory analysis. Scaling $\dot{V}_{O_2}$AT using estimated lean body mass was the only method studied that did not result in any 'spurious' correlations with body composition or body habitus. This is also depicted in Fig. \ref{fig:bc_scatter_VO2_bodycomp_elbm} on page \pageref{fig:bc_scatter_VO2_bodycomp_elbm}.}
\end{sidewaystable}