### 2016-03-14 response_to_reviewer’s_comment.md
Response letter in Markdown
- response.tex deleted
- new markdown file created

# Response to Reviewer’s comments

**An investigation of the clinical utility of preoperative cardiopulmonary exercise testing in patients undergoing major pancreatic surgery.**

Dear Sir,

I am grateful for the opportunity to respond to the examiners' comments. I have considered the examiners' comments carefully and these have improved the thesis. Please find my responses to the comments below.

## 1. Chapter 1 - More discussion of key RCTs

The following paragraph has been included as Section 1.x in Chapter 1.

## 2. Chapter 1 - Short discussion of aetiology needed

Point taken. The following subsection titled ’Aetiology of pancreatic cancer’ has been included in Chapter 1 under Section 1.1.
Several factors have been identified as being associated with an increased risk of pancreatic cancer. The most important risk factor associated with increased incidence of pancreatic cancer is tobacco. Obesity has also been reported to be associated with increased risk of pancreatic cancer although the underlying mechanisms are unclear. Other possible risk factors include increasing age, diet high in saturated fats, chronic pancreatitis, family history of pancreatic cancer and genetic abnormalities resulting in cancer syndromes. It is of note that some of the important risk factors such as smoking, obesity, poor diet and increasing age may also be associated with other medical comorbidities including cardiorespiratory diseases that may have a significant impact on patient fitness.

## 3. Chapter 2 - Review patient numbers on page 63

Point taken. Patient numbers have been updated to account for the missing information and this section now reads as below accounting for all the 100 patients included in this study.

Pathological examination of the resected specimen showed pancreatic ductal adenocarcinoma (n=37), ampullary adenocarcinoma (n=18), cholangiocarcinoma (n=17), duodenal adenocarcinoma (n=6), intra-ductal papillary mucinous neoplasia (n=4), neuroendocrine tumours (n=7), other neoplasia (n=6) or chronic pancreatitis (n=5).

## 4 Chapter 2 - Aim and end points must match

Point taken. The aims of this chapter have been updated as below to match the end-points described later in methods and results sections.

The aim of the present study was to evaluate the role of various measures of patient physiological fitness including cardiopulmonary exercise testing in predicting postoperative length of stay, major postoperative adverse events including operative mortality and fitness to undergo adjuvant therapy when indicated after pancreaticoduodenectomy.

## 5 Page 69 - Error in AT numbers

The results presented in this table have been re-analysed using the primary database and have been found to be accurate. In the group of patients where adjuvant chemotherapy was indicated, anaerobic threshold was low in a greater proportion. However, low anaerobic threshold was associated with non-receipt of adjuvant therapy in spite of being indicated as shown in this table.

# 6 Page 72 - Remove unjustified/unsupported conclusion, Page 73 - Remove GPS conclusion

I am not entirely sure how to address this. Should I completely remove this sentence?

Point taken. The following sentence has been removed from the discussion in Chapter 2.

Therefore, it would appear that objective measurement of patient physiological fitness using cardiopulmonary exercise testing is superior to conventional measures of patient fitness including the POSSUM Physiology Score or the modified Glasgow Prognostic Score and may have a role in predicting short-term outcome which in turn affects the overall management of these patients including receipt of adjuvant therapy.

## 7 Chapter 3, page 96 - Remove conclusions not based on thesis
Point taken. The following sentence has been removed from the discussion in Chapter 3.

Taken together, these results appear to suggest that preoperative biliary drainage and resolution of jaundice are unlikely to improve aerobic capacity and cardiopulmonary physiology in these patients.

## 8 Chapter 5,6 - Reduce to more understandable key issues. Several irrelevant tables should be removed.
Can we talk about this over the phone, please?

## 9 All acronyms/abbreviations on list
A list of abbreviations has been included.

## 10 All tables should have units (e.g. mg/L, etc) and legends.
Point taken. All tables have been updated to include units. Legends have been added to all figures and tables providing a brief description of the results
depicted.

## 11 All values and ranges need defining.
Are they looking for normal ranges for the various parameters like CRP,albumin, etc.?